# Learning Objectives {#r-rse-objectives}

```{r r-rse-objectives-setup, include=FALSE}
source("_common.R")
```

This appendix lays out the learning objectives for each set of lessons,
and is intended to help instructors who want to use this curriculum.

## The Basics of the Unix Shell
```{r, child="objectives/bash-basics.md"}
```

## Going Further with the Unix Shell
```{r, child="objectives/bash-advanced.md"}
```

## Git at the Command Line
```{r, child="objectives/git-cmdline.md"}
```

## Advanced Git
```{r, child="objectives/git-advanced.md"}
```

## Code Style, Review, and Refactoring
```{r, child="objectives/style.md"}
```

## Automating Analyses
```{r, child="objectives/automate.md"}
```

## Working in Teams
```{r, child="objectives/teams.md"}
```

## Project Structure
```{r, child="objectives/project.md"}
```

## Continuous Integration
```{r, child="objectives/ci.md"}
```

## R Packaging
```{r, child="objectives/package-r.md"}
```

## Python Packaging
```{r, child="objectives/package-py.md"}
```

## Correctness
```{r, child="objectives/correct.md"}
```

## Publishing
```{r, child="objectives/publish.md"}
```
