# Setting up an R package {#package-setup}

```{r setup, include=FALSE}
# TODO: Have this as its own setup chunk?
library(usethis)
```

## Learning Objectives

-   Identify when and why to create an R package
-   Learn how to create a new R package using
    `usethis::create_package()`
-   Describe and explain the purpose of the folders and files in a
    skeleton R package
-   Learn how and why you might add data to an R package
-   Describe what are global defaults in R and how to use them ease
    package development with `{devtools}` and `{usethis}`

## About Zipf's Law

TODO: Not sure this belongs here. Maybe in preface?

We'll create a package (`{zipfs}`) to analyze a fascinating result in
the field of quantitative linguistics. Zipf's Law states that the second
most common word in a body of text appears half as often as the most
common, the third most common appears a third as often, and so on. To
test whether Zipf's Law holds for a collection of classic novels that
are freely available from Project Gutenberg, we will write a software
package that counts and analyzes the word frequency distribution in any
arbitrary body of text.

## Why create an R package?

The main aim of developing software, and this applies to software
development in R as well, is to simplify some task or a set of tasks for
yourself and for others (like your team). And creating an R package is
one of the more powerful ways of simplifying tasks, no matter how simple
or complex to begin with.

TODO: Add a connecting paragraph or sentence...

How do you determine if you should make an R package? Ask these
questions to yourself. Have you ever made a function, no matter how
simple? Have you ever used that function more than once across more than
one project or file, either by copying and pasting it or by sourcing the
file that contains it? If you answered yes to these two questions, than
you should make an R package for that function. But you might wonder,
why make a package with only one function? Because it never stops with
*just* one function, you will very likely be making more in your work.
Aside from packaging up R functions, if you have small to medium sized
datasets that you use across multiple projects or files, you can even
package up data.

Aside from a "yes" to these questions, making an R package now is much
easier than it was in the past. That's largely thanks to R packages like
`{devtools}` and `{usethis}` that are designed to simplify developing R
packages. Combined with these packages, there are also so many free,
online books and resources for learning how to make R packages and doing
software development in R (LINK).

TODO: Include link to resources section for r packages book, mastering
software dev etc.

If we go back to the big picture in the preface in Section
@sec-big-picture, creating an R package (that is posted on GitHub or
submitted to CRAN) connected to scientific work or analyses will usually
fulfill the reproducibility and open science components of research.
That's because making an R package usually requires a higher level of
rigor, testing, documentation, and inspection of the code, which means
you and others will have more trust in the results. And if its on
GitHub, it's much easier to share your work with others and have a
bigger impact on research overall.

## What is an R package?

To understand what an R package is, we'll take a step back and consider
what happens when we interact with R. The very basic, low-level use of R
is through the R Console where you type functions in R to do tasks that
print to the screen. This initial R Console operates in your computer's
user home directory, usually something that looks like `/home/username/`
for Linux, `/Users/username/` for macOS, or `C:\Users\username` for
Windows. So any data or graphs that you save or import needs to include
the file path to the right location. From a reproducible and modular
point of view, this is goes against best practices. Likewise, the R code
entered into this session while in the Console is not
saved[^03-package-setup-1] So, you cannot easily re-use or share R code
when it is used this way.

[^03-package-setup-1]: Well, technically you can see the history by
    default in a file, but this stores *everything* you've entered into
    the R Console, including wrong code.

TODO: Image here that describes this a bit more?

The next level up is to save code in an R script (a file stored anywhere
on the computer that end in `.R`) and get R to run this code in
sequence. In the past, people might write R scripts in text editors like
vim, emacs, or Notepad, and manually run the script in the R Console.
Now, you'll mostly encounter people use an IDE (REF LINK) like RStudio
to do this work instead. This is the most commonly used approach to
doing any type of work in R. Theoretically, code written in this way is
reproducible and can be shared. However, practically, often R code in
these files is written out of sequence since researchers work with R
more interactively rather than programmatically. And the working
directory (REF LINK) of an R script changes depending on how it is used,
so we still encounter similar issues as if we used the R Console on its
own. You can use clever tricks to store commonly used R code or
functions in an R script and `source()` these files from other scripts,
among many other tricks. In essence, this involves keeping track of R
files across your computer to be able to do your work.

An R package is not much different from using several R scripts. There
are certain expectations and conventions that must be followed in order
for it to be "installed" on your computer as a R package. Some of these
conventions include:

-   There must be a file called `DESCRIPTION` what contains metadata
    necessary for R to know how to install the package. We'll cover what
    the fields should have in Section TODO: REF.

-   There must be an `R/` folder. This will usually contain only `.R`
    files and usually they contain only R functions you've created.

-   There must be a `NAMESPACE` file that contains the list of functions
    in your package that you want the user to use. We won't cover this
    file in detail because much of this file is managed automatically
    with the `{devtools}` and `{roxygen2}` packages.

-   The folder with the R code should be named the same as the package
    name. For instance, the R package usethis has the folder name of
    `{usethis}`. This is not an explicit requirement, but is highly
    recommended. R determines the name of the package from the
    `Package:` field in the `DESCRIPTION` file.

None of this is something you really have to worry about because the
packages `{usethis}` and `{devtools}` are designed to do many of these
package setup tasks easier for you.

So how does package installing work? When you tell R to install a
package, which is a set of files and folder that follow a specific
convention, R takes the folder and converts it into its own special
format. Then R saves this special format into a default location that is
specific to the operating system, like macOS, Windows, or Linux. To see
where packages are installed, use:

```{r}
.libPaths()
```

If you type this out on in your Console, what you see will probably be a
bit different to this. If `.libPaths()` outputs more than one path,
usually the first will contain all of the packages. Let's use the `{fs}`
package to explore this folder. First, let us see *some* of the packages
that are installed but using `dir_ls()`.

```{r}
library(fs)
# R priorities the first path, so we'll select it
primary_library_path <- .libPaths()[1]
packages <- dir_ls(path(primary_library_path))

# Number of installed packages
length(packages)
# See the first few package names
head(packages)
```

If we want to see how R stores packages in its special format, we can
use `dir_tree()` and look inside the installed `fs/` package folder.

```{r}
dir_tree(path(primary_library_path, "fs"))
```

See the `R/fs.rdb` and `R/fs.rdx` files? These are the special formats
that contain the functions within fs that R will look for when you load
the `{fs}` package. When a package is installed in R, what that means is
it is found in the R library folder given by `.libPaths()` with the
files in the R specific format. So when you run `library(fs)`, R will
know to search for the functions found in the `fs/` package folder.

So, installing your own package that is only found on your computer is
no different from installing a package from CRAN with
`install.packages()` or from GitHub with `remotes::install_github()`.
The only difference is that packages on CRAN or GitHub (from a public
repository) are usable to anyone who can access the Internet.

You might have noticed that we've used the syntax
`packagename::functionname()` a few times now. What this tells R to do
is to use the function from a specific package. So
`remotes::install_github()` is telling R to use the `install_github()`
function from the `{remotes}` package. For package development in
particular, we use this rather than, for instance, `library(remotes)`
because we want to be explicit about which function we are wanting to
use and because we don't need to load all the functions from the package
when we only want to use one or two of them.

### Exercise: Install a package from GitHub {#ex-install-github-package}

Since we make heavy use of `{usethis}` for building R packages, let's
get a bit familiar with it. Open up the [GitHub repository for
`{usethis}`](https://github.com/r-lib/usethis) and complete these tasks.

1.  Find which file contains the `create_package()` function in the
    [`R/`](https://github.com/r-lib/usethis/tree/main/R) folder. Hint,
    use GitHub's "Search or jump to..." feature to find things in the
    repository.

2.  Find out where the package name is in the `DESCRIPTION` file.

3.  In your RStudio Console, use `remotes::install_github()` to install
    the usethis package. View the help documentation
    (`?remotes::install_github`) to learn how to use it to install the
    package.

## Creating a package

Now that we have determined that *installed* R packages are found in the
library location determined by `.libPaths()`, what about packages that
you are actively developing? Do you save them in the same location?
Short answer, no, you should not. Instead create, develop, and save your
files somewhere in your Documents folder, or somewhere where you store
other work projects. Where exactly you save them is up to you, but as an
example, we might do them in `Documents/R/` or ... TODO: Add more
potential locations here.

Deciding where to save your package is the easy part. The hard part
comes with *naming* your package. Creating a meaningful, catchy, and
Google-able name is *really really hard*. There is no easy way to teach
you how to make one and it often requires several days of brainstorming
ideas. Once you do have an idea, you want to make sure that it hasn't
been taken already. You can do that easily with the package
`{available}`. We have already described in the preface (Section
\@ref(overview-packages-created)) that we will be creating a package
called `{zipfs}`, let's check to see if the package name is
`{available}`. The output will look a bit like this, though you'll see
more items.

```{r cs-available-pkg, eval=FALSE}
available::available("zipfs")
```

```{r ce-available-pkg, eval=TRUE, echo=FALSE}
available::valid_package_name("zipfs")
available::available_on_cran("zipfs")
available::available_on_bioc("zipfs")
available::available_on_github("zipfs")
```

TODO: Update based on if the package has been added to GitHub at time of
publication.

We see in this case that `{zipfs}` is available for us to use. Now that
we have a name, the next step is to setup the basic infrastructure for
the package. Thankfully, the `{usethis}` package helps ease a lot of the
difficult of this step. When we use the `usethis::create_package()`
function, it will setup the basic files and folders we need in order for
R to install it as a package. We'll setup use the function to create a
`{zipfs}` package.

TODO: Set author within GH Actions with usethis.full_name option. Use
local .Rprofile?

```{r cs-create-pkg, eval=FALSE}
usethis::create_package(path = "zipfs")
```

```{r ce-create-pkg, echo=FALSE}
# If needed. Code used in section, NOT shown to learners, but needed for auto building of package/book
temp_pkg <- fs::file_temp_push(glue::glue("{tempdir()}/zipfs"))
fs::dir_create(temp_pkg)
usethis::create_package(path = temp_pkg, open = FALSE)
```

Instead of using code, you can also do it through RStudio (if you use
it) by clicking through "File -\> New Project -\> ... TODO: Complete".
When you run this code in the Console, though, a bunch of text will pop
up (like above) indicating what was done on your computer. We'll go
through each of the things that have been created throughout the rest of
this section.

::: comment
TODO: Make sure this thing works for css and latex

You might notice that we are creating the package in the temporary
directory "`r fs::path_dir(temp_pkg)`". Because (almost) all of the
output from code in this book is generated automatically, we need to
have a way to create the package and do other code-based work that works
on each of the authors' computers as well as on the GitHub Actions
server (which builds the final version of this website). So we chose to
use the temporary location provided by R's `tempdir()` function. The
name of this temporary folder is random, so if you use it too, you will
see something different from what is in the book.
:::

As we wrote in the previous section, all R packages must have at least
an `R/` folder, a `DESCRIPTION` file, and a `NAMESPACE` file. Inside our
newly created package, we can see these have all been created. In
addition to the required files and folder, there are also the `.Rproj`
file since we use RStudio, the `.gitignore` file since we will
eventually use Git, and a `.Rbuildignore` file that tells R to ignore
including certain files when building the package. This will become
important later.

```{r ce-new-pkg-file-tree, echo=FALSE}
fs::dir_tree(temp_pkg, all = TRUE)
```

Open up the `DESCRIPTION` file so we can look inside it:

```{r ce-show-description-contents, results='asis', echo=FALSE}
# If needed. Code used in section, NOT shown to learners, but needed for auto building of package/book
cat(readLines(fs::path(temp_pkg, "DESCRIPTION")), sep = "\n")
```

As mentioned, the `DESCRIPTION` file contains the metadata for the
package and some configuration settings that R uses when building and
installing it. The format of the metadata uses the pattern `key: value`.
So for instance, the key `Package` has the value `zipfs`, which tells R
what the name of the package is. Likewise, `Title` and `Authors` give a
bit more information on what the package does and who created it. There
are many more keys to cover, some of which are not important to learn
about since they are created automatically and used internally by R for
general building purposes. But some are quite important. Specifically,
the `Title`, `Authors`, and `Description` keys should have more
appropriate values added to them.

### Exercise: Add details to the `DESCRIPTION` file {#ex-fill-in-description}

Let's fill in the important key fields in the `DESCRIPTION` file with
better details. If you haven't opened the `DESCRIPTION` file yet, please
open it up.

1.  In the `Title:` value field, replace the current generic text with
    with `Testing Zipfs Law on Novels`. Note the instructions to use one
    line only and to have the first letter of each word Capitalized.

2.  For the `Authors@R:` value field, go to the R Console and type out
    and run `?person`. Read through the help documentation and than fill
    it out as best you can in the `Authors@R:` section. The start should
    look like `Authors@R: person(`. For the `role` argument you'll need
    to add `c("cre", "aut")`. Read about what those mean in the
    `?person` help documentation. TODO: Not everyone may have this
    author key. Some might have `Authors:` or `Maintainer:`, need to
    consider how to handle this.

3.  In the `Description:` value field, replace the current generic text
    with a brief (one or two sentence) explanation

TODO: This should be as a solution maybe? The `DESCRIPTION` file should
look something like this now:

```{r ee-fill-in-description, include=FALSE}
# If needed. Code for exercise, NOT shown to learner, used for auto building of package/book
desc_file <- fs::path(temp_pkg, "DESCRIPTION")
desc <- desc::description$new(file = desc_file)
desc$set(
    Title = "Testing Zipfs Law on Novels",
    Description = "Contains functions to test out whether Zipfs Law applies to multiple novels."
)
# TODO: Decide on name of author to use through
desc$add_author(
    given = "Sami",
    email = "sami@random.com",
    role = c("aut", "cre")
)
desc$write()
```

```{r sol-fill-in-description}
# Code for exercise *solution*, NOT shown to learner, (maybe) used for auto building of package/book
desc
```

## Developer-friendly R settings

You will have noticed how often we use `::` when referring to functions
within packages. You will also probably have felt that it was a bit
annoying to always have to type these out, especially interactively in
the R Console. We will be using functions from the `{usethis}` and
`{devtools}` a lot as R package developers and for RSE work, so to make
our workflow easier, `{usethis}` has naturally made functions to help us
out. In the R Console, run each of these commands, following their
instructions before running the next one.

TODO: Include output from these?

```{r cs-developer-profile-options, eval=FALSE}
usethis::use_devtools()
usethis::use_usethis()
```

Don't forget to restart your session, using "Ctrl-Shift-F10" or "Session
-\> Restart R" or by using the "Command Palette" ("Ctrl-Shift-P" and
type out "Restart R") while using RStudio. After the R session restarts,
test out that the `{usethis}` and `{devtools}` packages are loaded by
typing out `use_` in the R Console and hitting the TAB key to list
available functions.

TODO: Add details about fixing issues if the above doesn't work.

After that, we want set up our RStudio so that whenever we open the R
Project of our `{zipfs}` package, the R session starts in a "blank
slate". What this means is that we don't want old data, functions, or
variables loaded into the Environment from the previous session we might
have worked on.

```{r cs-blank-slate-user, eval=FALSE}
use_blank_slate()
```

This function sets options to the user's Global Options in RStudio. If
you want these settings at a project level like the package so that when
you collaborate with others, everyone is using the same settings, use
instead:

```{r cs-blank-slate-project, eval=FALSE}
use_blank_slate("project")
```

```{r ce-blank-slate-project, echo=FALSE}
# If needed. Code used in section, NOT shown to learners, but needed for auto building of package/book
# local_project(temp_pkg, quiet = TRUE)
temp_pkg
usethis::with_project(temp_pkg, {
    use_blank_slate("project")
})
```

### Exercise: Setting up additional configurations

TODO: Should this be a note section or appendix or exercise?

There are other configurations you can set to make your work as a
developer easier, particularly when adding things to your `.Rprofile`
file. See the [usethis
vignette](https://usethis.r-lib.org/articles/usethis-setup.html) for
more details about it.

Open up your `.Rprofile` using `edit_r_profile()`. Inside copy and paste
the below text and replace the `___` with your specific values. If you
need help figuring out what to fill in for the `person()` section, use
`?person` to read through the help documentation. The last value about
the `Roxygen` and `markdown = TRUE` will be discussed more in the next
chapter.

``` r
options(
    usethis.full_name = "___",
    usethis.description = list(`Authors@R` = 'person(given = "___", family = "___", email = "___", role = c("aut", "cre"))',
                               Roxygen = "list(markdown = TRUE)")
)
```

## Package development workflow

While there are several ways of developing an R package, a
well-established and documented workflow revolves around making heavy
use of `{usethis}` and `{devtools}`. This is the workflow that we are
covering in this book and that we will briefly go over in this section.
Throughout this book you will be using this workflow many times, so
you'll get lots of practice with it.

The typical steps used in this development workflow are, in simple
terms, like this:

1.  When you have an idea for a function or set of similar function, for
    instance to load in text files, you would switch to the Console and
    type out `use_r("load-text")`. A new R file will be created in the
    `R/` folder and it will be opened up in RStudio.

2.  In this new file you would start creating and developing the
    function, writing the input arguments and the eventual output, while
    starting to fill in the Roxygen documentation. We'll discuss a more
    detailed function workflow in later chapters (TODO: add ref to
    chapters).

3.  Once you feel the function is ready to be tested out, you might test
    the function out by either running it in the `@examples` Roxygen
    section, by using a temporary "development" script or R Markdown
    file, or inside one of the R Markdown vignettes, all of which we
    will talk about more in later chapters (TODO: Add ref). To make the
    function available to R, you would use "Ctrl-Shift-L" while in
    RStudio to run `load_all()` (from `{devtools}`). After `load_all()`
    finishes, run the function to see if it works.

4.  Unless the function you created is very simple or in the *very*
    unlikely scenario that you happened to write code the worked
    perfectly the first time, the function will more likely not work
    exactly how you wanted or gives an error. In that case you go back
    to the function and continue developing it further, alternating
    between developing it, running `load_all()`, and testing the
    function out.

Let's try a bit of this workflow out by making a temporary, practice
function. In the Console, type out this function:

```{r cs-test-workflow, eval=FALSE}
use_r("test-workflow")
```

After it opens up, copy and paste the simple function below into the
newly created `R/test-workflow.R` file:

```{r cs-temp-function}
greeting <- function(name = "Sami") {
    paste0("Hi there, ", name, "!")
}
```

Before doing anything, type out `greeting()` in the Console and hit
enter to run it. What happens? Nothing! That's because R doesn't know
about the function just yet. We have to make R aware of it by either
running the function directly into R from the script (`Ctrl-Enter` to
send to the Console), by using `source()` on the script, by installing
the package, or by mimicking installing the package (but not actually
installing it) with the `load_all()` function. Since we're doing package
*development*, we will use the `load_all()` function, conveniently bound
to the keys `Ctrl-Shift-L` in RStudio.

Let us try out the `load_all()` function but hitting `Ctrl-Shift-L`.
After it finishes loading, again type out `greeting()` in the Console.
You should now see that it works.

### Exercise: Practice the workflow some more {#ex-practice-workflow-load-all}

Since you'll be using this workflow a lot when making R packages, we'll
do some more practices using it so you feel more comfortable with it.

Make the following changes to the function. With each change, re-load
(`Ctrl-Shift-L` in RStudio or `load_all()` in the Console) so that the
updated functions become available to R. After each re-load, switch to
the Console and run the `greeting()` function.

-   Change the default value for the `name` argument from `"Sami"` to
    `"Jun"`.

-   Change the greeting inside of the `paste0()` function from
    `"Hi there, "` to `"How do you do, "`.

-   Change the function `paste0()` to `paste()`. What changes?

After completing these tasks, since we don't need the
`R/test-workflow.R` file anymore, we'll delete it either manually or by
typing out:

```{r delete-test-workflow, eval=FALSE}
fs::file_delete("R/test-workflow.R")
```

## Installing instead of loading

When actively developing a package, rather than to directly use it in
another project, we use `load_all()`. That is because it can quickly
create an environment that mimics how the package might behave if it was
installed, without having to take the time to actually install it.

When we want to actually install the package, we use the `{devtools}`
function `install()` (keybinding is `Ctrl-Shift-B`), which makes the
package available to R on your computer. When you install, accessing the
functions in the package is done as you normally would with other
packages, like loading it with `library()`.

We won't cover this much more in this chapter, but for now, test it out
by hitting the `Ctrl-Shift-B` keys. Notice how much longer it takes
compared to `load_all()` (even if only a few seconds) and that the R
Session restarts, followed by a call to `library()`.

## Bundling data within a package

While you are able to bundle data within a package, the vast majority of
R packages do not contain data. In some instances, especially within
custom company or academic environments, packaging data in an R package
can make it very easy to share data with colleagues. And considering R
and many R packages are designed around working with data in some way,
like using statistics or making plots, we will also bundle data with our
package in order to make use of the data for the functions we will be
creating throughout this book.

### Starting with the raw data

Since Zipf's Law is about some math underlying text, the "data" we will
be working with is text from books. The books we are going to use come
from the open source [Project Gutenberg](https://www.gutenberg.org/).
Part of creating an R package is about making analytic or data
processing steps reproducible and explicit, so in this regard, since we
don't have the books we need on our computers, we need to write out
coding steps to get them. This is what the `data-raw/` folder is for
inside an R package. Inside this folder we include R scripts that have
code showing where and how we get the data that we do. To start, we will
set up the folder using the `{usethis}` function:

```{r cs-create-data-raw, eval=FALSE}
use_data_raw("project-gutenberg")
```

```{r ce-create-data-raw, echo=FALSE}
local_project(temp_pkg, quiet = TRUE)
use_data_raw("project-gutenberg", open = FALSE)
```

This function creates the `data-raw/` folder as well as an R script
inside, as we can see with `dir_tree()`.

```{r ce-show-tree-with-data-raw, echo=FALSE}
dir_tree(temp_pkg)
```

Inside the file is:

```{r cs-inside-empty-data-raw-folder, file=fs::path(temp_pkg, "data-raw", "project-gutenberg.R"), eval=FALSE}
```

We want to eventually download a few books but for now we will download
[Dracula](https://www.gutenberg.org/ebooks/345). Specifically, we want
the version of book that lets us load it into R, so the ["Plain Text
UTF-8"](https://www.gutenberg.org/cache/epub/345/pg345.txt) version is
the one we want. You find books by going onto [Project
Gutenberg](https://www.gutenberg.org/), typing out the book name in the
"Quick search" at the top left corner of the page, clicking the book
title and entering into the book page that should contain the link to
the plain text version. The steps we want to take are:

1.  Find a book, open the "Plain Text UTF-8" link, and copy the URL at
    the top.
2.  Using the URL, download the plain text book file and save it in the
    `data-raw/` folder. While we could read in the text directly from
    the URL, we want to keep a local copy in case we are working offline
    or in the unlikely chance that the website is down.
3.  Load the text file into R and save it into the `data/` folder. We'll
    explain why we do this step shortly.

Let's open up our `data-raw/project-gutenberg.R` script and complete the
exercise below to complete the steps above in R.

### Exercise: Download Dracula book and save as data

For this exercise, you will need to translate the steps listed above
into a form that R understands. To help break it down, what we need to
get R to do and write in the `data-raw/project-gutenberg.R` script is
to:

1.  Load the necessary packages: `{fs}`, `{readr}`, `{purrr}`, and
    `{here}`. We have so far only explained the `{fs}` package.
    -   `{readr}`: We use this to read in the text file.
    -   `{here}`: We use this for reproducibility purposes so that when
        we refer to file paths, they are always relative to the project
        root (where the `.Rproj` file is). Read the
        [website](https://here.r-lib.org/index.html) for more details on
        it.
    -   `{purrr}`: We use this to download and read in each book URL.
        Right now we only have one book, but we will be adding more
        later.
2.  Create a `list()` for the book's URL (`book_urls`), with the list
    item name being the books short title. While we are only downloading
    the Dracula book first, we want to write the code so that it is
    easier later to download more books later without modifying the code
    too much.
3.  Create a function (`download_then_read()`) that has arguments `url`
    and `book_name` that
    -   creates a file path from `book_name` to where the book should be
        downloaded to using `here()` pointing to `"data-raw"` and
        `paste0(book_name, ".txt")` (to add the text file ending)
    -   checks if the file exists (`file_exists()`) and downloads it
        (`download.file()`) if it does not
    -   then reads in the file into R (`read_lines()`) from `{readr}`
4.  Use `imap()` from `{purrr}`

```{r ee-for-copying-project-gutenberg-script-only, eval=FALSE, include=FALSE}
# Use for copying script code to add to chunk below.
readr::read_lines(here::here("data-raw/project-gutenberg.R")) |>
    clipr::write_clip()
```

```{r es-download-dracula-book, eval=FALSE}
library(____)
library(____)
library(____)
library(____)

# List with the short name for the book (no spaces or dashes/minuses), along
# the URL.
book_urls <- list(
    ___ = "___"
)

# Function to download the file if it doesn't exist, then read the file into R.
___ <- function(___, ___) {
    # Set the path of the book to `data-raw/` and ending as a `.txt` file.
    book_file_path <- ___("___", paste0(___, "___"))
    # Download text if it isn't already here
    if (!___(book_file_path)) {
        ___(url, book_file_path)
    }
    ___(book_file_path)
}

# `imap()` takes the URLs inside the `book_urls` list as the first argument (`url`) to
# `download_then_read()` and uses the `names()` of the list as the second
# argument (`book_name`). The output is a list of the books.
books <- ___(
    book_urls,
    download_then_read
)

# Save the book list as an .rda file in `data/`.
usethis::use_data(books, overwrite = TRUE)
```

```{r sol-project-gutenberg-script, file=here::here("data-raw", "project-gutenberg.R"), eval=FALSE, echo=-10}
```

### Loading the newly included data

-   Text: Data live in `data/` as `.rda` files.

    Restart, `devtools::load_all()`, `FIXME`

We'll cover how you will document data in Chapter
@sec-package-documentation.

### Exercise: Include the book Frankenstein in the books data

-   Exercise: Provide script for getting dracula text from Gutenberg.
    Add `dracula` to package.

1.  Summary of steps to do in exercise

```{r es-LABEL-NAME}
# Code for exercise and shown to learner
```

```{r ee-LABEL-NAME}
# If needed. Code for exercise, NOT shown to learner, used for auto building of package/book
```

```{r sol-LABEL-NAME}
# Code for exercise *solution*, NOT shown to learner, (maybe) used for auto building of package/book
```

## Final exercise

Complete these tasks as part of the package for your first project
assignment. Refer to the project assignment for exact details on what
you are required to create and on what the package should do. For this
exercise, we will be using a package about weather in Kenya *only as an
example and reference*.

1.  Decide on a simple but descriptive name for your new package. Use
    `{available}` to see if the name is taken.

2.  Use `create_package()` to create a new package. While you can create
    this package anywhere on your computer, for now we suggest either
    your `Desktop/` or maybe something like `Documents/r-pkgs/`.

3.  Open the `DESCRIPTION` file and update the file with:

    -   Yourself as the author in the `Authors@R:` field.
    -   A one sentence description of the package in the `Title:` field.
    -   Two or three sentence summary of the package in the
        `Description:` field.

4.  Add some data to the package. Use `use_data_raw()` to create the
    `data-raw/` folder along with a script for cleaning the raw data.
    Then put the data in this newly created folder. Write in some simple
    data cleaning processes in the cleaning R script and follow the
    steps in Section FIXME to make the FIXME data available as part of
    the package.

You'll know you are finished if:

-   Your package can be installed with `install()` (or `Ctrl-Shift-B` to
    build it).
-   In a fresh R session, after running `library(FIXME)` you can access
    the data by its name.

TODO: Add actions done to kenyaweather package here?

```{r ee-chapter-3-assignment, eval=FALSE}
# If needed. Code for exercise, NOT shown to learner, used for auto building of package/book
# TODO: Use temp or use Desktop or use a subfolder in this book's folder?
temp_dir <- fs::path_temp("kenyaweather")
usethis::create_package(
    temp_dir, 
    rstudio = TRUE, 
    open = FALSE,
    fields = list(
        # TODO: Decide which name to use
        `Authors@R` = 'person("Jane", "Doe", email = "jane@example.com",
                          role = c("aut", "cre"))',
        Title = "",
        Description = ""
    )
)

usethis::with_project(temp_dir, {
    usethis::use_blank_slate("project")
    # TODO: Change name for this.
    usethis::use_data_raw("weather", open = FALSE)
}, quiet = TRUE)

fs::dir_tree(temp_dir)
fs::dir_delete(temp_dir)
```

TODO: Bundle the final package used with Git/zip? move to temp folder?
TODO: Use git_commit/git_signature to set the author to Sami or other

