# Function development in a package environment

-   Process of creating non-function code, checking that it works, then
    converting it into a function
-   There are several workflows for this (create in vignette Rmd, make a
    dev/creating.R script as a development location, developing in the
    examples Roxygen section, 'Untitled1.R'). Which to use?
-   Mostly *how workflow actually looks like*.
-   Building functions up slowly, making small targeted functions that
    build up into a bigger more complex functions
-   Process control (if, stop)
-   Dependency management (better term? e.g. using `@importFrom` vs
    `::`)

TODO: should this be here or later?

-   Function documentation (with roxygen2), part 1
-   To use `@examples` to help with creating function

## Learning Objectives

-   Convert a chunk of R code into a function
-   Include a function in an R package
-   Provide basic function documentation using Roxygen
-   Use functions from other packages in a function in your own package
-   Return an error message from a function

## Introduction

So far, our zipfs package contains some data and a very simple function.
We've seen the general workflow for adding a function to a package, but
we skimmed over lots of the details. In this chapter, we'll take a much
closer look at writing functions, the details of including them in a
package, documenting them, and how to stop your functions with an error.

```{r ce-load-zipfs}
# Temporarily needed?, for getting access to `dracula`
library(zipfs)
```

## Writing Functions in R

We have a number of books included in our zipfs package. One useful
feature to add to our package would be a way to extract metadata about
the books from the book text. For example, let's take a look at lines 8
through 14 in Dracula:

```{r cs-dracula-meta}
load_all()
dracula[8:14]
```

    > load_all()
    ℹ Loading zipfs
    > dracula[8:14]
    [1] ""                                          
    [2] "Title: Dracula"                            
    [3] ""                                          
    [4] "Author: Bram Stoker"                       
    [5] ""                                          
    [6] "Release Date: August 16, 2013 [EBook #345]"
    [7] ""   

We can see the book text includes information on the book title, author
and release date. This format is consistent across the books, so it is
ideal for automatic parsing. Our goal will be to add a function
`book_meta()` that extracts these details from the text of a book.

Writing a function should start with working code. That is, rather than
starting with a blank function template for `book_meta()` and trying to
fill it in. First, we should write code that extracts the relevant
details in a concrete example, and then turn it into a function. We are
going to start with this working code that extracts these details from
the `dracula` object:

```{r cs-extract-meta, eval = FALSE}
library(stringr)
library(tibble)

devtools::load_all() # Replace with `library(zipfs)` if we've installed by now
text <- dracula[8:14]

author_line <- str_subset(text, "Author: ")
author <- str_remove(author_line, "Author: ")

title_line <- str_subset(text, "Title: ")
title <- str_remove(title_line, "Title: ")

release_date_line <- str_subset(text, "Release Date: ")
release_date <- str_remove(release_date_line, "Release Date: ")

tibble(
  author = author, 
  title = title,
  release_date = release_date
)
```

We are presenting this code fully formed, but usually getting to this
point is the hardest part of writing a function. We certainly didn't
write it out in one go. We started with a little example `text` then
wrote a little code, tested it out, fixed any problems and iterated.

But, where did we write this code? If we weren't working on a package,
we'd think nothing of opening up a new R script file, or Quarto file,
and putting our code inside, and saving it any old place. But, when we
are working on a package, we need to be more careful about adding files
that aren't directly related to the package. We wouldn't want someone
who wants to use our package also getting all code files that were
notes, or experiments, and in fact, there are rules about what files and
folders **can't** be in a package directory. One option would be to open
an R script and use it as a temporary code writing area, and never save
it - the `Untitled.R` strategy. This is fine for quick experiments, but
is a little dangerous for anything you might write over more than one
session. A better option is to create a specific location for code that
isn't part of the package, and make sure it is ignored when the package
is built. There is a `usethis` function `use_directory()` that helps set
this up. Let's create a `dev/` directory in our package project:

```{r cs-setup-dev-dir, eval = FALSE}
use_directory("dev", ignore = TRUE)
```

    ✓ Creating 'dev/'
    ✓ Adding '^dev$' to '.Rbuildignore'

The argument `ignore = TRUE` adds this folder to `.Rbuildignore` - a
file that lists things that shouldn't be included in the package when it
is built.

Now we can take the chunk above and save it in `dev/creating.R`, and be
sure it won't end up in the final built package. Let's work through the
code in that is in this chunk:

1.  Some packages are loaded - stringr is needed for the string
    operations and tibble for putting the results together.
2.  A variable `text` is created to hold a little bit of the `dracula`
    text as a concrete example of book text.
3.  The author is extracted from `text` into `author` by first pulling
    out the line that contains the pattern `"Author: "`, then removing
    the pattern `"Author: "` from that line.
4.  Repeat 3. but now with the pattern `"Title:  "` to extract the
    title.
5.  Repeat 3. but now with the pattern `"Release Date:  "` to extract
    the release date.
6.  Combine `author`, `title`, and `release_date` in a one row tibble.

If we run the entire script we should end with the output:

    # A tibble: 1 × 3
      author      title   release_date                
      <chr>       <chr>   <chr>                       
    1 Bram Stoker Dracula August 16, 2013 [EBook #345]

### When to write a function?

-   Identify repetition in above code: x3

```{=html}
<!-- -->
```
    line <- str_subset(text, XXX)
    value <- str_remove(line, XXX)

-   Review how to turn code into a function: pick a name, identify
    arguments, fill in the function constructor.

```{r cs-extract-element}
extract_element <- function(text, element){ 
  line <- str_subset(text, element)
  value <- str_remove(line, element)
  value
}
extract_element(text, "Author: ")
```

-   Explicit vs implicit returns?

```{r cs-implicit-return}
extract_element <- function(text, element){
  line <- str_subset(text, element)
  str_remove(line, element)
}
extract_element <- function(text, element){
  line <- str_subset(text, element)
  value <- str_remove(line, element)
  return(value)
}
```

-   Updated chunk

```{r cs-updated-chunk}
library(stringr)
library(tibble)

text <- dracula[8:14]

tibble(
  author = extract_element(text, "Author: "), 
  title =  extract_element(text, "Title: "),
  release_date = extract_element(text, "Release Date: ")
)
```

### Exercise: Write `book_meta()` {#ex-write-book-meta}

Write the `book_meta()` function.

This is how we'd like our chunk to look:

```{r es-book-meta}
library(stringr)
library(tibble)

text <- dracula[8:14]
book_meta(text)
```

```{r sol-book-meta}
book_meta <- function(text){
  tibble(
    author = extract_element(text, "Author: "), 
    title =  extract_element(text, "Title: "),
    release_date = extract_element(text, "Release Date: ")
  )
}
```

## Including Functions in an R Package

Current contents of `dev/creating.R`

```{r cs-creating-contents}
library(stringr)
library(tibble)
load_all() # I don't think we've `install()`ed yet?

text <- dracula[8:14]

extract_element <- function(text, element){ 
  line <- str_subset(text, element)
  value <- str_remove(line, element)
  value
}

book_meta <- function(text){
  tibble(
    author = extract_element(text, "Author: "), 
    title =  extract_element(text, "Title: "),
    release_date = extract_element(text, "Release Date: ")
  )
}

book_meta(text)
```

-   Text: Functions live in `.R` files in `R/`. How should functions be
    organized into files?

```{r ce-create-code-file}
use_r("book-meta")
```

-   Text: move function definitions into `R/book-meta`

```{r cs-contents-book-meta}
extract_element <- function(text, element){ 
  line <- str_subset(text, element)
  value <- str_remove(line, element)
  value
}

book_meta <- function(text){
  tibble(
    author = extract_element(text, "Author: "), 
    title =  extract_element(text, "Title: "),
    release_date = extract_element(text, "Release Date: ")
  )
}
```

-   Text: Now, `load_all()` loads our function definitions too.

```{r ce-updated-creating-contents}
library(stringr)
library(tibble)
load_all() # I don't think we've `install()`ed yet?
book_meta(dracula) # notice this is now all lines 
```

### Exercise: Write `count_words()` based on chunk

Might put code in `dev/counting.R`

```{r es-count-words-chunk}
library(stringr)
library(tibble)
library(zipfs)

# convert to all lower
text_lower <- str_to_lower(dracula) 

# split by one or more whitespace characters
chunks <- unlist(str_split(text_lower, "\\s+"))

# remove punctuation
words <- str_remove_all(chunks, pattern = "[[:punct:]]+")

# get rid of empty strings
words <- words[words != ""]

# count occurrences and sort
word_freq <- sort(table(words), decreasing = TRUE)

# return in rectangular format
tibble(word = names(word_freq), n = as.numeric(word_freq))
```

Reminder of steps:

-   Convert chunk into function in `dev/counting.R`. Check it works
    `count_words(dracula)`
-   Create new file in `R/` called `count-words.R`, copy over definition
-   Check work by editing code in `dev/counting.R` to

```{r es-check-count}
library(stringr)
library(tibble)
library(zipfs)

count_words(dracula)
```

```{r sol-count-words-chunk}
count_words <- function(text) {
    # convert to all lower
    text_lower <- str_to_lower(text) 
    
    # split by one or more whitespace characters
    chunks <- unlist(str_split(text_lower, "\\s+"))
    
    # remove punctuation
    words <- str_remove_all(chunks, pattern = "[[:punct:]]+")
    
    # get rid of empty strings
    words <- words[words != ""]
    
    # count occurrences and sort
    word_freq <- sort(table(words), decreasing = TRUE)
    
    # return in rectangular format
    tibble(word = names(word_freq), n = as.numeric(word_freq))
}
```

## Dependencies: Using Functions From Other Packages

-   Text: We have a problem. On clean session, `load_all()`,
    `book_meta(dracula)`, produces an error, can't find `tibble()`.
    Difference between interactive work (`library(tibble)`) and package
    development (don't want to load entire libraries in user's
    workspaces)

```{r ce-dep-error, error = TRUE}
# Restart R, clean slate
load_all()
book_meta(dracula)
```

-   Text: Two changes needed: specify our package relies on another,
    inside functions, be specific about where functions inside our own
    code come from.

-   Text: Specify our function relies on another package,

```{r ce-use-package}
use_package("tibble")
```

-   Text: What changes? Adds line to Imports in DESCRIPTION, ensures
    when someone installs our package, they get tibble too if they don't
    have it.

Output includes:

    • Refer to functions with `tibble::fun()`

Text: Also need to be explicit when we use tibble functions that these
come from the tibble package, use `tibble::` in front

`book_meta()` becomes:

``` {cs-book-meta-tibble}
book_meta <- function(text){
  tibble::tibble(
    author = extract_element(text, "Author: "), 
    title =  extract_element(text, "Title: "),
    release_date = extract_element(text, "Release Date: ")
  )
}
```

Text: Try again, restart R,

```{r ce-dep-error-2}
# Restart R, clean slate
load_all()
book_meta(dracula)
```

Text: Still have an error, but now it's related to functions we use
inside `extract_element()`.

### Exercise: Fix dependency problems in `extract_element()`

Reminder of steps:

-   Add the stringr package as a dependency with
    `usethis::use_package()`
-   Anywhere a stringr function is used preface with `stringr::`

You'll know it's worked if this code works without error in a fresh
session of R.

```{r es-extract-element-dep}
# Restart R, clean slate
load_all()
book_meta(dracula)
```

```{r sol-extract-element-dep}
# + change to DESCRIPTION
extract_element <- function(text, element){ 
  line <- stringr::str_subset(text, element)
  value <- stringr::str_remove(line, element)
  value
}
```

### Exercise: Fix dependency problems in `count_words()`

<!-- Doesn't use pipe anymore, so this is just finding the stringr and tibble functions and adding stringr:: and tibble:: -->

## Documenting Functions

-   Text: Outline of Roxygen workflow: special comments above
    functions + `document()`, generates `man/` and allows
    `?function_name`, Maybe explain all this using an existing package.

-   Code: With `book_meta()` RStudio: Code -\> Insert ROxygen Skeleton,
    `document()`, `?book_meta()`

``` {cs-book-meta-roxygen-skeleton}
#' Title
#'
#' @param text 
#'
#' @return
#' @export
#'
#' @examples
book_meta <- function(text){
  tibble::tibble(
    author = extract_element(text, "Author: "), 
    title =  extract_element(text, "Title: "),
    release_date = extract_element(text, "Release Date: ")
  )
}
```

``` {cs-document}
document()
?book_meta
```

-   Text: Common Roxygen tags (title, arguments, return, examples), demo
    with `book_meta()`

``` {cs-book-meta-roxygen-complete}
#' Get book metadata from a Gutenberg book
#'
#' @param text A character vector of lines from a Gutenberg book
#'
#' @return a one row tibble with `author`, `title` and `release` columns.
#' @export
#'
#' @examples
#' book_meta(dracula[1:20])
book_meta <- function(text){
  tibble::tibble(
    author = extract_element(text, "Author: "), 
    title =  extract_element(text, "Title: "),
    release_date = extract_element(text, "Release Date: ")
  )
}
```

### Exercise: Document `count_words()`

Reminder of steps:

1.  Insert a roxygen skeleton above the definition of `count_words()`
2.  Run `document()` to generate help page and preview with
    `?count_words`
3.  Edit the roxygen comment.
4.  Iterate steps 2. and 3. until done.

```{r sol-document-count-words}
#' Count word frequency in text
#'
#' Counts the occurrence of words over all elements provided in `text`.
#'
#' Each element of `text` is considered a line.
#' Lines are first converted to lower case then split into words
#' by whitespace. Then punctuation is removed, and empty strings are ignored.
#'
#' @param text A character vector
#'
#' @return A tibble with columns `word`,
#' and `n` the number of times `word` occurs in `text`.
#' @export
#'
#' @examples
#' count_words(dracula)
```

-   Text: `@export`, why we might not document every function., e.g.
    `extract_element()`

## Stopping Your Function with an Error

-   Text: Alternate workflow for editing functions, use `@examples`,
    edit directly in `R/book-meta.R`. Add `book_meta(1)` as example.

```{r cs-add-example}
#' Get book metadata from a Gutenberg book
#'
#' @param text A character vector of lines from a Gutenberg book
#'
#' @return a one row tibble with `author`, `title` and `release` columns.
#' @export
#'
#' @examples
#' book_meta(dracula[1:20])
#' book_meta(1)
```

<!-- This will cause a check error later, because examples need to run without error. -->

-   Text: Basic checks of user input, how to return an error

Basic template of input check:

``` r
if( << test for bad input >> ){
    stop("<< message for bad input >>", call. = FALSE)
}
```

-   Text: As a comment? Mention other kinds of process control, give
    pointer to learn more.

```{r ce-bad-input-ideas}
text <- 1
is.numeric(text) 
is.character(text)
!is.character(text)
```

Guidance for writing error messages:
<https://style.tidyverse.org/error-messages.html>

See if you want fancier bulleted list type error messages:
<https://rlang.r-lib.org/reference/abort.html>

```{r ce-bad-input-check}
if(!is.character(text)){
    stop("`text` must be a character vector.", call. = FALSE)
}
```

Add to function definition

```{r cs-book-meta-with-check, eval = FALSE}
book_meta <- function(text){
    if(!is.character(text)){
        stop("`text` must be a character vector.", call. = FALSE)
    }
    tibble::tibble(
        author = extract_element(text, "Author: "), 
        title =  extract_element(text, "Title: "),
        release_date = extract_element(text, "Release Date: ")
    )
}
```

`load_all()` Rerun examples

### Exercise: Add an input check to `count_words()`.

```{r sol-input-check-count-words}
count_words <- function(text) {
    if(!is.character(text)){
        stop("`text` must be a character vector.", call. = FALSE)
    }
    
    # convert to all lower
    text_lower <- str_to_lower(text) 
    
    # split by one or more whitespace characters
    chunks <- unlist(str_split(text_lower, "\\s+"))
    
    # remove punctuation
    words <- str_remove_all(chunks, pattern = "[[:punct:]]+")
    
    # get rid of empty strings
    words <- words[words != ""]
    
    # count occurrences and sort
    word_freq <- sort(table(words), decreasing = TRUE)
    
    # return in rectangular format
    tibble(word = names(word_freq), n = as.numeric(word_freq))
}
```

## Wrapping Up

-   Text: Re-iterate idea of extracting common code into simpler
    functions. Suggest input check could be its own function. Then
    describe steps as a review of chapter: decide on a name, decide on a
    file to store it, write it, try it, careful with dependencies,
    document it.

## Final exercise {#04-final-exercise}

TODO: INCOMPLETE, need to think about this more.

Complete these tasks as part of the package for your first project
assignment. The exact functions you should create will depend heavily on
what your package will do. Get help and guidance from your instructor.
Use our examples on the `kenyaweather` package as further guidance.

1.  Create a new R script using `use_r("..")`

2.  Add a `FIXME()` function to your FIXME package

    The following code snippet produces ... FIXME.

    ```{r, eval = FALSE}
    # A line or two of R code that does something useful
    FIXME
    ```

    Turn it into a function called `FIXME()` that takes arguments for
    `FIXME` and `FIXME` and include it in your FIXME package.

    Insert a Roxygen Skeleton for this function, making sure to edit the
    title, description, argument descriptions, and example. Don't forget
    to run `document()` to produce the help page.

3.  Add a `FIXME2()` function to your FIXME package

    Add another function called `FIMXE2()`, that ... FIXME. It should
    return an error if FIXME. *This function is intentionlly given
    without a snippet of code so you can practice the function
    development workflow.*

    Insert a Roxygen Skeleton for this function, making sure to edit the
    title, description, argument descriptions, and example. Don't forget
    to run `document()` to produce the help page.

While you are completing these tasks, You'll know you are done when
after installing your package and starting a fresh R session:

-   The following code runs without error and produces the right output

    ```{r, eval = FALSE}
    library(FIXME)
    FIXME()
    FIXME2()

    # Should give an error
    FIXME2()
    ```

-   `?FIXME()` and `?FIXME2()` bring up help pages that have at least a
    title, a description of each argument, and at least one example of
    usage.

TODO: Add actions done to kenyaweather package here?

## Final exercise

<!-- NOTE: Preferably something linked to working towards completing the project assignments, maybe to be used as a milestone? -->
