# Function development in a package environment

- Process of creating non-function code, checking that it works, then
   converting it into a function
- There are several workflows for this (create in vignette Rmd, make a dev/creating.R
    script as a development location, developing in the examples Roxygen
    section, 'Untitled1.R'). Which to use?
- Mostly *how workflow actually looks like*.
- Building functions up slowly, making small targeted functions that build up
  into a bigger more complex functions
- Process control (if-else, stop, return, switch)
- Dependency management

TODO: should this be here or later?

- Function documentation (with roxygen2), part 1
- To use `@examples` to help with creating function

## Final exercise {#04-final-exercise}

TODO: INCOMPLETE, need to think about this more.

Complete these tasks as part of the package for your first project assignment.
The exact functions you should create will depend heavily on what your package
will do. Get help and guidance from your instructor. Use our examples on the 
`kenyaweather` package as further guidance.

1. Create a new R script using `use_r("..")`

1. Add a `FIXME()` function to your FIXME package
    
    The following code snippet produces ... FIXME.  
    
    ```{r, eval = FALSE}
    # A line or two of R code that does something useful
    FIXME
    ```
    
    Turn it into a function called `FIXME()` that takes arguments for `FIXME`
    and `FIXME` and include it in your FIXME package.
    
    Insert a Roxygen Skeleton for this function, making sure to edit the title,
    description, argument descriptions, and example. Don't forget to run
    `document()` to produce the help page.

2. Add a `FIXME2()` function to your FIXME package

    Add another function called `FIMXE2()`, that ... FIXME. It should return an
    error if FIXME. *This function is intentionlly given without a snippet of
    code so you can practice the function development workflow.*
    
    Insert a Roxygen Skeleton for this function, making sure to edit the title,
    description, argument descriptions, and example. Don't forget to run
    `document()` to produce the help page.

While you are completing these tasks, You'll know you are done when after
installing your package and starting a fresh R session:

* The following code runs without error and produces the right output

    ```{r, eval = FALSE}
    library(FIXME)
    FIXME()
    FIXME2()
    
    # Should give an error
    FIXME2()
    ```
    
* `?FIXME()` and `?FIXME2()` bring up help pages that have at least a title, a description of each argument, and at least one example of usage.

TODO: Add actions done to kenyaweather package here?
