# Developing documentation and tutorials on usage {#package-documentation}

<!-- python book documentation: https://merely-useful.tech/py-rse/scripting.html#scripting-docstrings-->

```{r cs-usethis, include=FALSE}
library(usethis)
```

## Learning Objectives {#learning-objectives-8}

-   Describe the purpose of documenting code and packages
-   Document functions with `{roxygen2}`
-   Create a README and vignette for package documentation
-   Ensure that documentation is consistent with up-to-date style
    guidelines using `{styler}`

## Why do we need documentation? {#introduction}

Just like version control tracks what changes you make and why,
documentation for code and packages tells your collaborators, your
package's users, and future you *what* the code does and *why*.

Over time, programming languages have become more "human-readable": it
is possible for an experienced programmer to read code and understand
what it does. But even if one understands *what* a chunk of code is
doing, it may not be obvious at all *why* it does that! Programming
languages come with built-in ways to add comments and documentation that
are not read by the computer, but exist solely to help you and your
collaborators describe and understand code. Good documentation is key
for openness, reproducibility, and project sustainability.

Good documentation happens at all scales of code: at the level of
individual objects, functions, and tests as well as for the entire
package. In this chapter, we describe the workflow of documentation
alongside package development and outline the components of a
well-documented package.

## Documenting functions and tests with roxygen2 {#function-documentation}

In R, documentation files for objects are stored by default in the
`man/` folder in your package directory with a `.Rd` extension. These
help files can be accessed for an object, function, or package by typing
`?object-name` or `help(object-name)`. Writing documentation for
functions as you write the function itself is helpful for future you and
other users to understand how to use your code. You can edit these
documentation files directly, but the package `{roxygen2}` provides some
shortcuts to make it easier to generate the help files with standard
formatting.

Let's add documentation for the function `extract_element()` from the
zipfs package. First, all we need to do is write the function itself.

```{r cs-extract-element-fun}
extract_element <- function(text, element){
  line <- str_subset(text, element)
  str_remove(line, element)
}
```

Next, with your cursor over the function definition, click
`Code > Insert roxygen skeleton` in RStudio (or use the command palette
with `Ctrl-Shift-P` and then type out `roxygen` to search for the
option). This command adds the default components of function
documentation as comments above the function.

```{r cs-extract-element-with-roxygen}
#' Title
#'
#' @param text 
#' @param element 
#'
#' @return
#' @export
#'
#' @examples
extract_element <- function(text, element){
  line <- str_subset(text, element)
  str_remove(line, element)
}
```

We can edit the default template to add more information in each field.
The fields `@return` and `@export` are optional, depending on the
function's behavior. We'll also add an example of how to use the
function.

<!-- TODO: perhaps include a deliberate typo here to be caught later by `spelling`?-->

```{r cs-extract-element-fill-in-roxygen}
#' Extract specific text element from string.
#'
#' @param text A string of text
#' @param element A substring to search for in `text`
#'
#' @return The matching line without the text matched by `element`
#' @export
#'
#' @examples
#' text <- "Title: Dracula"
#' extract_element(text, "Title: ")
extract_element <- function(text, element){
  line <- str_subset(text, element)
  str_remove(line, element)
}
```

Next, press `Ctrl/Cmd + Shift + D` or run `devtools::document()` in the
Console to generate or update a file named `man/extract_element.Rd`. Try
typing `?extract_element()` in the console to see what happens! (You may
need to run `devtools::load_all()` in order to test out the new help
file.)

<!-- TODO: I wasn't able to get the man file to generate - 'Error: Clipboard on X11 requires that the DISPLAY envvar be configured.' -->

### Exercise: Add roxygen comments to functions in zipfs {#ex-roxygen}

<!-- TODO: All functions in the draft package have documentation except `dpower` and `nllpower` - we can document these ones, or take out other documentation to do this exercise. -->

Follow the steps above to add documentation for the `dpower()` and
`nllpower()` functions in the zipfs package. What files are created in
the `man/` folder when multiple functions in the same file are
documented?

### Documenting tests

TODO: Decide if this is needed based on what is put into chapter 6.

-   Code: FIXME test documentation workflow
-   Exercise: add and/or improve comments and documentation in a test
    file in the zipfs package.

Reference: ([R Packages
reference](https://r-pkgs.org/man.html?q=roxygen#man))

## The README: The first stop documentation shop {#create-readme}

<!-- Some bits from py-rse/packaging -->

When a user or contributor first encounters a package, they usually want
to know what the package is meant to do, instructions on how to install
it, and examples of how to use it. We can include these elements in the
README file.

The function `usethis::use_readme_rmd()` creates an R Markdown document
with some boilerplate sections and text. These default "Installation"
and "Example" sections are a great starting point to create a helpful
and descriptive README.

```{r cs-readme, eval=TRUE}
use_readme_rmd()
```

### Exercise: R Markdown vs. Mardown {#ex-rmd-md}

Read the help file for `use_readme_rmd()` and `use_readme_md()`. What is
the difference between `.Rmd` and `.md` files? Why might one or the
other be better for a README?

GitHub treats the file `README.md` specially, displaying it at the top
of your repo for easy access. The README created by `use_readme_rmd()`
contains the basics of what a package README should have.

```{r ce-show-readme-contents, echo=FALSE, results='asis'}
# This prints the contents as output in the book.
cat(usethis:::render_template(
    "package-README",
    data = list(on_github = TRUE, Package = "zipfs"),
    package = "usethis"
), sep = "\n")
```

### Knitting the Markdown README {#knit-readme}

To create the Markdown README that GitHub will display, you can use
`{knitr}` to knit the RMD file into `README.md`. The YAML text
`output: github_document` included in the template tells `{knitr}` to
create a Markdown output for GitHub instead of the default HTML. In
RStudio, click `Knit` in the menu above the editor, or run
`devtools::build_readme()` in the Console as suggested by the template
text.

The main difference between `devtools::build_readme()` and knitting is
that `build_readme()` will do some additional checks that will be
important for submitting to CRAN. This step is really only important
before you submit to CRAN, not while actively developing the package.

### Exercise: Edit and knit the zipfs README {#ex-readme}

Edit `README.Rmd` to include a description of zipfs, installation
instructions, and an example of its use. Knit the `README.Rmd` file so
it creates a `README.md` file.

## Vignettes: Long-form documentation of examples {#vignette}

Everything we've discussed about documentation so far applies to all
programming projects in nearly any language: All programming languages
provide some way to document code, and all GitHub projects have READMEs.
"Vignettes" are a distinctly R type of documentation - they describe a
problem-solving unit of your package in detail, with examples.

The function `browseVignettes()` allows you to, you guessed it, browse
existing vignettes. Calling `browseVignettes()` with no arguments will
open a browser window with a list of vignettes for all installed
packages, while `browseVignettes("packagename")` will show vignettes for
a particular package. You can also call up a vignette directly if you
know its name with `vignette("vignettename")`.

```{r cs-vignette, eval=FALSE}
browseVignettes("tidyr")
vignette("nest")
```

### Exercise: Finding package vignettes {#ex-find-vignette}

Find the vignette(s) for the packages `{devtools}` and `{ggplot2}`. Why
might a package have more than one vignette? In what ways is a vignette
different from a README?

### Making a vignette {#make-vignette}

Once again, `{usethis}` has a useful function to help us create
vignettes. The function `usethis::use_vignette()` takes a package name
and vignette title and creates a short template R Markdown document in
the `vignettes/` folder.

```{r cs-create-vignette, eval=TRUE}
use_vignette("zipfs", title = "Using zipfs")`
```

The boilerplate text is very minimal. Try knitting the file
`vignette/zipfs.Rmd` to see what's included by default - it's not much.
Vignettes can vary a lot depending on the goals of the author; they can
be long or short,
[simple](https://cran.r-project.org/web/packages/devtools/vignettes/dependencies.html)
or
[complex](https://cran.r-project.org/web/packages/ggplot2/doc/extending-ggplot2.html).

### Exercise: Edit a zipfs vignette {#ex-edit-vignette}

<!-- This might be too involved for a single exercise -->

Modify the vignette `zipfs.Rmd` (in the `vignettes/` folder) to include
a one-paragraph description of the purpose of zipfs, and an example of
how to use the zipfs functions together to analyze a book. Knit the file
to preview the vignette.

Reference: [R packages](https://r-pkgs.org/vignettes.html)

## Writing in style: The housekeeping help of styler and spelling {#styler-spelling}

When writing code, some aspects of formatting are very strict, but there
are still choices a programmer can make about the appearance and content
of code. For instance, R strictly requires that parentheses be properly
paired in order for code to function (`print("hi")` will run,
`print("hi"` will not), but things like variable names and the position
of whitespace are more flexible. To create code that is maximally
understandable, many communities have developed *style guides* that
outline guidance for how to write clear code and documentation. Check
out the [tidyverse style guide](https://style.tidyverse.org/)'s
[syntax](https://style.tidyverse.org/syntax.html) section for examples
of the types of stylistic decisions we make when writing code.

Different communities may disagree on certain aspects of style.
[Google's R style
guide](https://google.github.io/styleguide/Rguide.html) lists
differences between their guide and the tidyverse style guide. For
instance, tidyverse recommends that all object names be written in
`snake_case` (lowercase with underscores), while Google prefers function
names to be written in `BigCamelCase` (capitalizing the first letter of
each word). Which style guide you adhere to will depend on who your
project is for, who you're working with, and on your own personal
preferences.

It's possible to adhere to a style guide by remembering the rules and
applying them yourself, but luckily there are tools that can help
automate the boring parts. The package `{styler}` automatically formats
code according to a specified style guide: Simply run
`styler::style_pkg()` to format your code according to the tidyverse
style guide by default.

In the same spirit, the package `{spelling}` does a spell check on the
documentation files in a package. Run `spelling::spell_check_package()`
prints a list of potential spelling errors and their locations to the
console; unlike `{styler}`, it does not modify files.

### Exercise: Checking style {#ex-styler}

Run `style_pkg()` to restyle the files in the zipfs package. Use Git to
explore some of the changes that were made. What do you notice?

### Exercise: Checking spelling {#ex-spelling}

Run `spell_check_package()` to check the spelling in the zipfs
documentation files. Fix the errors, then run `update_wordlist()` to add
any special words to the spell-checker that you want to ignore.

## Final exercise {#08-final-exercise}

Complete these tasks as part of the package for your first project
assignment. Refer to the project assignment TODO: Add ref to assignment
here? for more details. As with the other final exercises, we will use
the kenyaweather package as an example and reference. Don't forget to
add and commit any changes you've made into the Git history after each
task, and pushing to GitHub.

1.  Add the ability to write your roxygen2 documentation using Markdown
    with `usethis::use_roxygen_md()`.

2.  Go through your existing function documentations and fill them in
    more. Especially focus on completing the `@param`, `@return`, and
    `@examples` fields. Then run `document()` (`Ctrl-Shift-D`).

3.  Create a README file for your repository by using
    `use_readme_rmd()`.

    -   Edit the `README.Rmd` to include a description of your package
        and its purpose, installation instructions, and one or more
        examples of how to use it.
    -   Knit the `README.Rmd` file so it creates a `README.md` file.

4.  Add a vignette to your project package by running
    `use_vignette("YOURPACKAGENAME")`, replacing `YOURPACKAGENAME` with
    the name of your actual project package. Edit the vignette to
    include a:

    -   Higher-level paragraph describing your package and its purpose.
    -   An example of how to use your package, including code, that's
        more complete than what the README might have.
    -   Use `knitr` to Knit and preview your vignette.

5.  (*Alternate*) Add a vignette to your project that describes how your
    package interacts with another published R package. <!--leaflet?-->

    -   Use `{knitr}` to Knit and preview your vignette.

6.  Check the spelling and style of your files in your package by
    running `styler::style_pkg()` to restyle your files and
    `use_spell_check()` to check the spelling. Fix any problems and if
    there are any special words, create a new file in `inst/WORDLIST`,
    add the word, and run `spelling::update_wordlist()`. Then re-run
    these styling and spelling checks to confirm no problems exist
    anymore.
