# Involving the community {#community-involvement}

```{r, cs-usethis, eval=TRUE, echo=FALSE}
library(usethis)
```

## Learning objectives {#learning-objectives-10}

- Edit a project README and NEWS file to be welcoming to contributors and users
- Understand the importance of having an explicit Code of Conduct and how to
create one relevant to a project
- Understand the purpose of licensing content, choose a license and create a
LICENSE file
- Understand the importance of contributing guidelines and how to develop them
for a project
- Create and manage GitHub Issues to track project bugs and improvements

## Creating a welcoming community with welcoming materials {#welcoming-materials}

Now that your project is taking shape
and you even have a website to share information about it,
there are things you can do to help other people understand
how to interact with your project.
This chapter covers the information users expect to find if they would like to use 
and perhaps contribute to your project. 

### Creating a welcoming README {#readme}

The README is the first part of your project that new users and contributors see,
and by making it clear and welcoming, you make it possible for people from
many different fields or levels of experience to understand what you're doing and why.

You created a README in Chapter 5, and in Chapter 8 you filled out the
basic building blocks of a README: a description of your project, instructions
for installing and using your project, and some examples. 
Now you will return to your README and assess it from the perspective of a new
user or contributor. Is it welcoming? Is it easy to understand? Does it have
all the information needed to use or contribute to your project?

We recommend paying special attention to the following components
to make your README clear and welcoming.
For an example, check out the [README from the STEMM Role Models project](https://github.com/KirstieJane/STEMMRoleModels#stemm-role-models-app).

- A great way to start a welcoming README is with a statement of welcome! You
can introduce yourself, welcome people to your project, and say that you're
excited they came.

- Write your project description to answer the following questions in a way that is
understandable to a wide range of people. Avoid using jargon, and if you
do use terms or concepts that are field-specific, define them.
    - What is this project about?
    - Who is this project for?
    - Why does this project exist?
    - What makes your project unique or exciting?

- Describe how to use your project with simple examples. More on this in the
[reprex](#reprex) section.

- Describe what kinds of help or resources you are looking for. More on this in the
[contributing guidelines](#contributing-guidelines) section.

Remember: even if you are pretty sure that all your users have the same expertise
as you or that you won't have contributors, writing a clear README is also
extremely helpful for your number one collaborator: Future You. Furthermore, the very
act of thinking through your project's documentation can be a great way to
clarify your own thoughts, brainstorm ideas, and foresee issues. 

References: [Mozilla Open Leadership Training Series](
https://mozilla.github.io/open-leadership-training-series/articles/opening-your-project/write-a-great-project-readme/)

### Exercise: Write a jargon-free description of the `zipfs` package. {#ex-upgoer5}

When you are an expert in some particular subject matter, it can be hard
to know which of the words and concepts you use often are not so easy
for others to understand. This exercise will challenge you to rethink
how you explain concepts by taking it to the extreme - restricting your vocabulary
to the 1000 most common words in the English language!

Write a one-paragraph description of the `zipfs` package using
the [Up-Goer Five Text Editor](https://splasho.com/upgoer5/).
<!--- For example: "These computer words tell the computer to look at a book and add up the number of times each word is used and then make a picture of all the numbers. Most of the time, all the numbers turn out to be a straight line."--->

### Using reproducible examples to explain your project {#reprex}

https://community.rstudio.com/t/faq-whats-a-reproducible-example-reprex-and-how-do-i-create-one/5219

Simple examples are very effective tools for explaining how your package works.
The `reprex` package is designed to help you create **rep**roducible **ex**amples:
stand-alone examples that include all the necessary information about running
a chunk of code in the same way someone else did. This packages is mainly aimed
at creating bug reports, but it is also a great tool for packaging examples
more generally. 

Let's create a reprex for how to use the `count_words` function in the `zipfs` package
by following these steps:
- Install (only for the first time) and load the `reprex` package.

```{r cs-reprex-load, eval=FALSE}
install.packages('reprex')
library(reprex)
```

- Highlight the entire chunk of code that you want to include in your reprex, 
including any packages, and copy the text using Ctrl/Cmd + c.

FIXME: should this be from the perspective of developing zipfs, like starting with `load_all()` instead of `library(zipfs)`?

```{r cs-reprex, eval=TRUE}
library(zipfs)
text <- c("Star light, star bright,
    First star I see tonight;
    I wish I may, I wish I might,
    Have the wish I wish tonight.")
(word_freq <- count_words(text))
```

- In the console, type `reprex()` and hit Enter/Return.
- Your reprex is now stored on your clipboard - use Ctrl/Cmd + v to paste it where you like.

Reference: [So you’ve been asked to make a reprex](https://www.jessemaegan.com/blog/so-you-ve-been-asked-to-make-a-reprex/)

### Exercise: use `reprex` to create a minimal reproducible example for the `book-meta` function {#ex-reprex}

Create a reprex for simple usage of the `book-meta` function in the `zipfs` package.
FIXME: this might be tricky since it requires data - need to test this to make sure it's doable, or 
talk about how to include data?

### Keeping current with a NEWS file {#news}

When you make changes to your project, such as fixing a bug or adding
a new feature, it's a good idea to maintain a high-level summary of
these changes so that people who interact with your project can understand
when and how your project is changing. By convention, this list of
changes lives in a file called NEWS.md. Changes
are usually grouped by which package release they belong to; we will
cover releasing your package in more detail in Chapter 12. Check out
the [devtools NEWS.md](https://github.com/r-lib/devtools/blob/main/NEWS.md)
for an example of how this file is used in a popular package. 

The function `usethis::use_news_()` automatically creates a file called
NEWS.md with a header and the start of a bullet-point list.
Let's run it now for the `zipfs` package. 

```{r cs-news, eval=TRUE}
use_news_md()
```

Each time you make a change,
add it to the NEWS.md file as you go. When it's time for a release,
you'll have a handy summary of all the changes that have been made.

[reference](https://r-pkgs.org/release.html#news))

### Exercise: edit the NEWS.md file {#ex-news}

Make an edit to `NEWS.md` in the `zipfs` package describing the
most recent change that was make to the structure or code of `zipfs` (FIXME: we
can be more specific here, like referring to a change made in the previous chapter.)

### Contributing guidelines {#contributing-guidelines}

By this point, your README already contains some information about
how people can use or get involved in your project. When you want
to include more specific details, those details will live in a file called
CONTRIBUTING.md. Contributing guidelines help people understand how
to interact with your project and help you plan out the types of
contributions you are looking for. We'll talk more in a later chapter
(\@ref(team-package-development)) about how to develop contributing guidelines
for working with a team of developers.

The function `usethis::use_tidy_contributing()` automatically creates a file called
CONTRIBUTING.md in the folder `.github` with some boilerplate contributing
guidelines in the Tidyverse style.

```{r, cs-contributing, eval=TRUE}
use_tidy_contributing()
```

The specific topics and text in your contributing guidelines will depend
a lot on your project. For example, you might want to add sections about
reporting bugs, how to communicate with others on the project,
or where to find more resources or information. Don't forget to make your
include a statement welcoming people to your project at the top!
Here's the [GitHub Docs CONTRIBUTING.md file](https://github.com/github/docs/blob/main/CONTRIBUTING.md)
as an example. 

[reference](https://mozilla.github.io/open-leadership-training-series/articles/building-communities-of-contributors/write-contributor-guidelines/)

### Exercise: Reading contributing guidlines {#ex-contributing}

Choose two open source products that you have used and find their
contributing guidelines. Do they contain any sections that aren't in the
default Tidyverse Contributing Guidelines that you think are useful?

### Code of Conduct {#code-of-conduct}

To create an inclusive and safe community of users and contributors,
a Code of Conduct is an important document to include in your project.
A Code of Conduct improves safety by clearly laying out which behaviors
are encouraged and which are not acceptable, helping to prevent problems
from arising in the first place, as well as outlining how problems
will be handled if they do occur. 

The function `usethis::use_code_of_conduct()` creates a file called
`CODE_OF_CONDUCT.md` that contains the text of the 
[Contributor Covenant](https://www.contributor-covenant.org/).

```{r, cs-code_of_conduct, eval=TRUE}
use_code_of_conduct()
```

The files CONTRIBUTING.md and CODE_OF_CONDUCT.md are treated specially
by GitHub so that when a new issue is opened for the first time,
a message box appears with a link to CONTRIBUTING.md and
CODE_OF_CONDUCT.md, directing new contributors to read these documents.

### Exercise: Editing the Code of Conduct {#ex-code-of-conduct}

Read the Code of Conduct file in the `zipfs` package and edit it
to include a method of contact for reporting violations. Add details for how
violations will be handled (for example, banning from project, blocking, etc.).

### Exercise: Developing a Code of Conduct as a group {#ex-group-code-of-conduct}

Discuss as a group what your expectations are for working together
(things like communication channels, communication expectations,
timeliness, meeting preparedness, meeting etiquette, etc.).

While your code of conduct list in this exercise may
include things that are specific to your group, the code of conduct templates
that are widely used (such as the tidyverse default) have been tested across
many environments and for many different people. We recommend that you add to
your code of conduct as needed to suit your situation, but that you keep the
standard components to create the most welcoming and safe environment possible. 

[reference](https://mozilla.github.io/open-leadership-training-series/articles/building-communities-of-contributors/write-a-code-of-conduct/)

### Licensing {#license}

While a Code of Conduct describes how contributors should interact with each other,
a license dictates how project materials can be used and redistributed.
If the license or a publication agreement makes it difficult for people to contribute,
the project is less likely to attract new members,\index{project!license}
so the choice of license is crucial to the project's long-term sustainability.

> **Open Except...**
>
> Projects that are only developing software may not have any problem making everything open.
> Teams working with sensitive data, on the other hand,
> must be careful to ensure that what should be private isn't inadvertently shared.
> In particular,
> people who are new to Git (and even people who aren't)
> occasionally add raw data files containing personal identifying information to repositories.
> It's possible to rewrite the project's history to remove things when this happens,
> but that doesn't automatically erase copies people may have in forked repositories.
Every creative work has some sort of license;
the only question is whether authors and users know what it is and choose to enforce it.
Choosing a license for a project can be complex,
not least because the law hasn't kept up with everyday practice.
@Mori2012 and @Vand2014 are good starting points
to understand licensing and intellectual property from a researcher's point of view,
while @Lind2008 is a deeper dive for those who want details.
Depending on country, institution, and job role,
most creative works are automatically eligible for intellectual property protection.
However,
members of the team may have different levels of copyright protection.
For example,
students and faculty may have a copyright on the research work they produce,
but university staff members may not,
since their employment agreement may state that
what they create on the job belongs to their employer.

To avoid legal messiness,
every project should include an explicit license.
This license should be chosen early,
since changing a license can be complicated.
For example,
each collaborator may hold copyright on their work
and therefore need to be asked for approval when a license is changed.
Similarly,
changing a license does not change it retroactively,
so different users may wind up operating under different licensing structures.

> **Leave It to the Professionals**
>
> Don't write your own license.
> Legalese is a highly technical language,
> and words don't mean what you think they do.

To make license selection for code as easy as possible,
GitHub allows us to select one of several common software licenses when creating a repository.
The Open Source Initiative maintains [a list][osi-license-list] of \gref{open licenses}{open_license},
and [choosealicense.com][choose-license] will help us find a license that suits our needs.
Some of the things we need to think about are:

1.  Do we want to license the work at all?
2.  Is the content we are licensing source code?
3.  Do we require people distributing derivative works to also distribute their code?
4.  Do we want to address patent rights?
5.  Is our license compatible with the licenses of the software we depend on?
6.  Do our institutions have any policies that may overrule our choices?
7.  Are there any copyright experts within our institution who can assist us?

Unfortunately,
GitHub's list does not include common licenses for data or written works like papers and reports.
Those can be added in manually,
but it's often hard to understand the interactions among multiple licenses
on different kinds of material [@Alme2017].

Just as the project's Code of Conduct is usually placed in a root-level file called `CONDUCT.md`,
its license is usually put in a file called `LICENSE.md`\index{project files!LICENSE}
that is also in the project's root directory.

### Exercise: The Landscape of Open Source Licenses {#ex-license}

1. Explore [choosealicense.com](https://choosealicense.com/) and read
about the different types of open source licenses.

2. Choose two open source products that you have used (for example
[Firefox](https://www.mozilla.org/en-US/MPL/), 
[Atom](https://github.com/atom/atom/blob/master/LICENSE.md), or 
[Wordpress](https://wordpress.org/about/license/)). Find and read the license
for each product.

#### Software {#teams-license-software}

In order to choose the right license for our software,
we need to understand the difference between two kinds of license.\index{open source license}
The \gref{MIT License}{mit_license}\index{MIT License}\index{open source license!MIT License}
(and its close sibling the BSD License)\index{BSD License}\index{open source license!BSD License}
say that people can do whatever they want to with the software as long as they cite the original source,
and that the authors accept no responsibility if things go wrong.
The \gref{GNU Public License}{gpl} (GPL)\index{GNU Public License (GPL)}\index{GPL (GNU Public License)}\index{open source license!GPL}
gives people similar rights,
but requires them to share their own work on the same terms:

> You may copy, distribute and modify the software as long as you track changes/dates in source files.
> Any modifications to or software including (via compiler) GPL-licensed code
> must also be made available under the GPL
> along with build and install instructions.
>
> --- [tl;dr][tldr-gpl]
In other words,
if someone modifies GPL-licensed software or incorporates it into their own project,
and then distributes what they have created,
they have to distribute the source code for their own work as well.

The GPL was created to prevent companies from taking advantage of open software
without contributing anything back.
The last thirty years have shown that this restriction isn't necessary:
many projects have survived and thrived without this safeguard.
We therefore recommend that projects choose the MIT license,
as it places the fewest restrictions on future action.

```text
MIT License
Copyright (c) 2020 Amira Khan
Permission is hereby granted, free of charge, to any person
obtaining a copy of this software and associated documentation
files (the "Software"), to deal in the Software without
restriction, including without limitation the rights to use,
copy, modify, merge, publish, distribute, sublicense, and/or
sell copies of the Software, and to permit persons to whom the
Software is furnished to do so, subject to the following
conditions:
The above copyright notice and this permission notice shall be
included in all copies or substantial portions of the Software.
THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES
OF MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND
NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT
HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY,
WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING
FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR
OTHER DEALINGS IN THE SOFTWARE.
```

> **First, Do No Harm**
>
> The [Hippocratic License][hippocratic-license]\index{Hippocratic License}\index{open source license!Hippocratic License}
> is a newer license
> that is quickly becoming popular.
> Where the GPL requires people to share their work,
> the Hippocratic License requires them to do no harm.
> More precisely,
> it forbids people from using the software in ways that violate
> the [Universal Declaration of Human Rights][udhr].
> We have learned the hard way that software and science can be mis-used;
> adopting the Hippocratic License is a small step toward preventing this.

#### Data and reports {#teams-license-other}

The MIT license, the GPL, and the Hippocratic License are intended for use with software.
When it comes to data and reports,
the most widely used family of licenses are those produced
by [Creative Commons][creative-commons].\index{Creative Commons}
These have been written and checked by lawyers and are well understood by the community.

The most liberal option is referred to as \gref{CC-0}{cc_license},\index{Creative Commons!CC-0 license}\index{public domain}
where the "0" stands for "zero restrictions."
This puts work in the public domain,
i.e.,
allows anyone who wants to use it to do so, however they want, with no restrictions.
CC-0 is usually the best choice for data,
since it simplifies aggregate analysis involving datasets from different sources.
It does not negate the scholarly tradition and requirement of citing sources;
it just doesn't make it a legal requirement.

The next step up from CC-0 is the Creative Commons--Attribution license,
usually referred to as \gref{CC-BY}{cc_license}.\index{Creative Commons!CC-BY license}\index{CC-BY license}
This allows people to do whatever they want to with the work
as long as they cite the original source.
This is the best license to use for manuscripts:
we want people to share them widely
but also want to get credit for our work.

Other Creative Commons licenses incorporate various restrictions,
and are usually referred to using two-letter abbreviations:

-   ND (no derivative works) prevents people from creating modified versions of our work.
    Unfortunately, this also inhibits translation and reformatting.

-   SA (share-alike) requires people to share work that incorporates ours
    on the same terms that we used.
    Again,
    it is fine in principle, but in practice it makes aggregation and recombination difficult.

-   NC (no commercial use) does *not* mean that people cannot charge money for something that includes our work,
    though some publishers still try to imply that in order to scare people away from open licensing.
    Instead,
    the NC clause means that people cannot charge for something that uses our work without our explicit permission,
    which we can give under whatever terms we want.

To apply these concepts to our Zipf's Law project,
we need to consider both our data (which other people created)
and our results (which we create).
We can view the license for the novels by looking in `data/README.md`,
which tells us that the Gutenberg Project books are in the public domain (i.e., CC-0).
This is a good choice for our results as well,
but after reflection,
we decide to choose CC-BY for our papers
so that everyone can read them (and cite them).

The function `usethis::use_NAME_license()`, replacing `NAME` with a license from the list, will create a file called LICENSE.md with the text of that license. Hint: start typing `use_license` in RStudio to see a dropdown of the different licenses that are included in the `usethis` package.

```{r, cs-license, eval=TRUE}
use_ccby_license()
```

## Project management for a user-friendly package

Codes of conduct and licenses are a project's constitution,
but how do contributors know what they should actually be doing on any given day?
Whether we are working by ourselves or with a group of people,
the best way to manage this is to use an \gref{issue tracking system}{issue_tracking_system}
to keep track of tasks we need to complete or problems we need to fix.
\gref{Issues}{issue} are sometimes called \gref{tickets}{ticket},\index{project!issue}\index{issue (in project)}
so issue tracking systems are sometimes called \gref{ticketing systems}{ticketing_system}.\index{issue tracking system}
They are also often called \gref{bug trackers}{bug_tracker},
but they can be used to manage any kind of work,
and are often a convenient way to manage discussions as well.

\newpage

Like other \gref{forges}{forge},
GitHub allows participants to create issues for a project,
comment on existing issues,
and search all available issues.
Every issue can hold:

-   A unique ID, such as `#123`, which is also part of its URL.
    This makes issues easy to find and refer to:
    GitHub automatically turns the expression `#123` in a \gref{commit message}{commit_message}
    into a link to that issue.

-   A one-line title to aid browsing and search.

-   The issue's current status.
    In simple systems (like GitHub's) each issue is either open or closed,
    and by default,
    only open issues are displayed.
    Closed items are generally removed from default interfaces,
    so issues should only be closed when they no longer require any attention.

-   The user ID of the issue's creator.
    Just as `#123` refers to a particular issue,
    `@name` is automatically translated into a link to that person.
    The IDs of people who have commented on it or modified it are embedded in the issue's history,
    which helps figure out who to talk to about what.

-   The user ID of the person assigned to review the issue, if someone is assigned.

-   A full description that may include screenshots,
    error messages,
    and anything else that can be put in a web page.

-   Replies, counter-replies, and so on from people who are interested in the issue.

Broadly speaking,
people create three kinds of issues:

1.  \gref{Bug reports}{bug_report} to describe problems they have encountered.\index{bug report}

2.  \gref{Feature requests}{feature_request}\index{feature request}
    describing what could be done next,
    such as "add this function to this package" or "add a menu to the website."

3.  Questions about how to use the software,
    how parts of the project work,
    or its future directions.
    These can eventually turn into bug reports or feature requests,
    and can often be recycled as documentation.

> **Helping Users Find Information**
>
> Many projects encourage people to ask questions on a mailing list or in a chat channel.
> However,
> answers given there can be hard to find later,
> which leads to the same questions coming up over and over again.
> If people can be persuaded to ask questions by filing issues,
> and to respond to issues of this kind,
> then the project's old issues become a customized [Stack Overflow][stack-overflow] for the project.
> Some projects go so far as to create a page of links
> to old questions and answers that are particularly helpful.
## Bug Reports {#teams-bugs}

One steady supply of work in any active project is \gref{bug reports}{bug_report}.
Unsurprisingly,
a well-written bug report is more likely to get a fast response\index{bug report!style}
that actually addresses the problem [@Bett2008].
To write a good bug report:

1.  Make sure the problem actually *is* a bug.
    It's always possible that we have called a function the wrong way
    or done an analysis using the wrong configuration file.
    If we take a minute to double-check,
    or ask someone else on our team to check our logic,
    we could well fix the problem ourselves.

2.  Try to come up with a \gref{reproducible example}{reprex}
    or "reprex"\index{reproducible example (reprex)}\index{reprex (reproducible example)}
    that includes only the steps needed to make the problem happen,
    and that (if possible) uses simplified data rather than a complete dataset.
    Again,
    we can often solve the problem ourselves as we trim down the steps to create one.

3.  Write a one-line title for the issue
    and a longer (but still brief) description that includes relevant details.

4.  Attach any screenshots that show the problem,
    resulting errors,
    or (slimmed-down) input files needed to re-create it.

5.  Describe the version of the software we were using,
    the operating system we were running on,
    which version of the programming language we ran it with,
    and anything else that might affect behavior.
    If the software in question uses a logging framework (Section \@ref(errors-logging)),
    turn debugging output on and include it with the issue.

6.  Describe each problem separately so that each one can be tackled on its own.
    This parallels the rule about creating a branch in version control for each bug fix or feature
    discussed in Section \@ref(git-advanced).

An example of a well-written bug report with all of the components mentioned above
is shown in Figure \@ref(fig:teams-bug-report).

```{r teams-bug-report, echo=FALSE, fig.cap="Example of a bug report."}
insert_graphic("figures/teams/bug-report.png")
```

It takes time and energy to write a good error report.
If the report is being filed by a member of the development team,
the incentive to document errors well is that resolving the issue later is easier.
You can encourage users from outside the project to write thorough error reports
by including an issue template for your project.
An issue template is a file included in your GitHub repository
that proliferates each new issue with text that describes expectations
for content that should be submitted.
You can't force new issues to be as complete as you might like,
but you can use an issue template to make it easier for contributors to
remember and complete documentation about bug reports.

Sometimes the person creating the issue may not know or have the right answer for some of these things,
and will be doing their best with limited information about the error.
Responding with kindness and encouragement is important to maintain a healthy community,
and should be enforced by the project's Code of Conduct\index{Code of Conduct (for project)}
(Section \@ref(teams-coc)).

## Labeling Issues {#teams-labels}

The bigger or older a project gets,
the harder it is to find things---unless, that is,
the project's members put in a bit of work to make things findable [@Lin2020].
Issue trackers let project members add \gref{labels}{issue_label} to issues\index{issue (in project)!label}
to make things easier to search and organize.
Labels are also often called \gref{tags}{tag};\index{tags (for issues in project)}
whatever term is used,
each one is just a descriptive word or two.

GitHub allows project owners to define any labels they want.
A small project should always use some variation on these three:

-   *Bug*: something should work but doesn't.

-   *Enhancement*: something that someone wants added to the software.

-   *Task*: something needs to be done, but won't show up in code
    (e.g., organizing the next team meeting).

Projects also often use:

-   *Question*: where is something
    or how is something supposed to work?
    As noted above,
    issues with this label can often be recycled as documentation.

-   *Discussion* or *Proposal*: something the team needs to make a decision about
    or a concrete proposal to resolve such a discussion.
    All issues can have discussion:
    this category is for issues that start that way.
    (Issues that are initially labeled *Question*
    are often relabeled *Discussion* or *Proposal*
    after some back and forth.)

-   *Suitable for Newcomer* or *Beginner-Friendly*:
    to identify an easy starting point for someone who has just joined the project.
    If we help potential new contributors find places to start,
    they are more likely to do so [@Stei2014].

The labels listed above identify the kind of work an issue describes.
A separate set of labels can be used to indicate the state of an issue:

-   *Urgent*: work needs to be done right away.
    (This label is typically reserved for security fixes).

-   *Current*: this issue is included in the current round of work.

-   *Next*: this issue is (probably) going to be included in the next round.

-   *Eventually*: someone has looked at the issue and believes it needs to be tackled,
    but there's no immediate plan to do it.

-   *Won't Fix*: someone has decided that the issue isn't going to be addressed,
    either because it's out of scope or because it's not actually a bug.
    Once an issue has been marked this way,
    it is usually then closed.
    When this happens,
    send the issue's creator a note explaining why the issue won't be addressed
    and encourage them to continue working with the project.

-   *Duplicate*: this issue is a duplicate of one that's already in the system.
    Issues marked this way are usually also then closed;
    this is another opportunity to encourage people to stay involved.

Some of the labels that GitHub creates for repositories by default
are shown in Figure \@ref(fig:teams-labels).
These labels can be modified or otherwise customized for each repository.

```{r teams-labels, echo=FALSE, fig.cap="Labels for GitHub issues."}
insert_graphic("figures/teams/issue-labels.png")
```

Some projects use labels corresponding to upcoming software releases, journal issues, or conferences
instead of *Current*, *Next*, and *Eventually*.
This approach works well in the short term,
but becomes unwieldy as labels with names like `sprint-2020-08-01` and `spring-2020-08-16` pile up.

Instead,
a project team will usually create a \gref{milestone}{milestone},\index{milestone (in project)}\index{project!milestone}
which is a set of issues and pull requests in a single project repository.
GitHub milestones can have a due date
and display aggregate progress toward completion,
so the team can easily see when work is due and how much is left to be done.
Teams can also create projects,
which can include issues and pull requests from several repositories
as well as notes and reminders for miscellaneous tasks.

--------------

Your package now lives in a GitHub repo, and you've created several helpful files that will orient users and contributors to your project. In this section, we discuss some of GitHub's features that are useful for managing your project.

GitHub Issues provide a way to log bugs and project improvement suggestions
from you, collaborators, and your user community. The Issues tab on GitHub stores issues in reverse chronological order and sorts them into an 'Open' and 'Closed' category by default. Creating [issue templates](
https://docs.github.com/en/communities/using-templates-to-encourage-useful-issues-and-pull-requests/configuring-issue-templates-for-your-repository) for bugs and feature requests.

  - Text: creating, using, and managing GitHub Issues to maintain your project.
GitHub Issues provides a way to log bugs and project improvement suggestions
from you, collaborators, and your user community. Creating [issue templates](
https://docs.github.com/en/communities/using-templates-to-encourage-useful-issues-and-pull-requests/configuring-issue-templates-for-your-repository) for bugs and feature requests.
  - Exercise: create an issue template for reporting bugs in the `zipfs` package.
  - Exercise: create an issue describing a possible new feature or enhancement
for the `zipfs` package.
  - Text: project maintenance, longevity and sustainability. Possible areas to
cover: [project maintenance](
https://mozilla.github.io/open-leadership-training-series/articles/open-project-maintenance/open-project-maintenance/),
contributor [personas and pathways](https://mozilla.github.io/open-leadership-training-series/articles/building-communities-of-contributors/bring-on-contributors-using-personas-and-pathways/),
[leadership structure](https://mozilla.github.io/open-leadership-training-series/articles/building-communities-of-contributors/understand-meaningful-participation-and-distributed-leadership/),
[event planning and facilitation](https://mozilla.github.io/open-leadership-training-series/articles/running-awesome-community-events/event-planning-and-facilitation/).
  
### Tracking tasks with GitHub Issues (moved from chapter 12)
  - Text: Beyond bugs: how GitHub issues can help you keep track of tasks and
communicate with your community ([reference](
https://mozilla.github.io/open-leadership-training-series/articles/get-your-project-online/project-set-up-for-collaboration-with-github/#project-structure-and-organization)). 
How to label issues and link issues to other parts of your project.
  - Exercise: create at least three issue labels for the `zipfs` repository.
Label the existing issues so that each has at least one label.
  - We'll discuss using labels for project management in \@ref(team-package-development)

Also moved from chapter 12:
- use_github_labels(), browse_github_issues(), issue_reprex_needed(), and perhaps issue_close_community()
- Give set of recommended labels?
- Issue templates? (made by Joel: https://github.com/altair-viz/altair/issues/new/choose)

This chapter lays the foundation for encouraging other people to interact with your code.
In following chapters, we'll formalize the the methods for sustaining your project.

<!---
NOTE: This is potential content to use, if it fits.

And some people still worry that if they make their data and code generally available,
someone else will use it and publish a result they have come up with themselves.
This is almost unheard of in practice,
but that doesn't stop people using it as a scare tactic.
Other people are afraid of looking foolish or incompetent by sharing code that might contain bugs.
This isn't just [impostor syndrome][impostor-syndrome]:
members of marginalized groups are frequently judged more harshly than others,
so being wrong in public is much riskier for them.
-->


## Final exercise {#10-final-exercise}

Complete these tasks as part of the package for your first project assignment.
Refer to the project assignment TODO: Add ref to assignment here?
for more details. As with the other final exercises,
we will use the `kenyaweather` package as an example and reference.
Don't forget to add and commit any changes you've made into the Git history
after each task, and pushing to GitHub. For more information and tips at any
step, refer to the 
[Mozilla Open Leadership Training Series resources](https://mozilla.github.io/open-leadership-training-series/articles/readme/)

1. Review your project package's `README.Rmd`:
    - Paste your package description into the
    [Up-Goer Five Text Editor](https://splasho.com/upgoer5/).
    Try to revise it to include only allowed words. *Note*: You don't
    need to use this in your actual README, but it is a fun and informative
    exercise to see and understand what words you think are common, but might not be!
    - Review your `README.Rmd` file by trying to place yourself with the
    viewpoint of a new contributor to your project. Does your README clearly
    explain the purpose and function of your package without using jargon? 
    Is it welcoming and clearly laid out? Make
    as many additions, deletions, and revisions as you'd like until you feel
    your README is clear and welcoming.
    - After editing your `README.Rmd` file, make sure to knit the `README.md`
    version so it gets properly displayed on GitHub.
 
2. Add a Code of Conduct to your package repository:
    - Create a file in your package called `CODE_OF_CONDUCT.md` by running
    `usethis::use_code_of_conduct()`. This function automatically populates your
    Code of Conduct with the
    [Contributor Covenant](https://www.contributor-covenant.org/). 
    The newly created file will be in the `.github/` folder. 
    You may keep or modify the template, or you may draft your own entirely.
    Make sure to attribute the source of any outside text you use or modify, and
    make sure the license of the text you use permits your use or modification.
    Here are some other example codes of conduct:
        - [Citizen Code of Conduct](https://github.com/stumpsyn/policies/blob/master/citizen_code_of_conduct.md)
        - [Django Code of Conduct](https://www.djangoproject.com/conduct/)
    - Include a link to your Code of Conduct in your `README.Rmd`. Knit the
    file so it updates the `README.md` version.
 
3. Create a contributing guidelines file for your package repository:
    - Create a file called `CONTRIBUTING.md` by running
    `usethis::use_tidy_contributing()`. 
    The newly created file will be in the `.github/` folder. 
    This file will contain information for
    people interacting with your project about how they can use your package
    and/or how they can get involved. Review the newly created file and
    revise it as necessary. For some examples of great Contributing
    Guidelines, check out:
        - [Atom](https://github.com/atom/atom/blob/master/CONTRIBUTING.md)
        - [Data Carpentry Geospatial R lesson](https://github.com/datacarpentry/r-intro-geospatial/blob/master/CONTRIBUTING.md)
        - [Tidyverse](https://www.tidyverse.org/contribute/)
    - The default `.github/CONTRIBUTING.md` file you created has a template for
    instructions for submitting changes or reporting a problem. Modify
    the `CONTRIBUTING.md` file to include the following:
        - A welcome statement and a brief description of your project.
        - A list of project-related resources, like a link to your README.
        - A list of communication channels your project uses, if applicable
        (mailing lists, forums, chat clients, GitHub Issues, etc.)
        TODO: The above might be more appropriate for the project management
        section.
        - Instructions for how to submit changes or report a problem, specific
        to your project.
        - A description of how you recognize contributors.
        - Instructions on where to go for help: who to contact and how.
        - A link to your Code of Conduct.
 
4. Add a license to your repository for terms of re-use of your package:
    - Choose a license for your package. Visit
    [choosealicense.com](https://choosealicense.com/) and choose a license that
    makes sense for your project. This license will live in a file called
    `LICENSE.md` in the top level folder of your project. If you choose a common
    license, it's likely that you can create `LICENSE.md` by running
    `usethis::use_mit_license()` (for the MIT License) or another related
    `usethis::use_NAME_license()` function. Licensing is a complex thing
    so for now, choose the MIT License.

5. Project management
- Create an [issue template](https://docs.github.com/en/communities/using-templates-to-encourage-useful-issues-and-pull-requests/configuring-issue-templates-for-your-repository) 
for your project, providing guidelines for what you think might be the most
common category of issue that contributors and users might submit.
- Add a line to CONTRIBUTING.md requesting that new issues use the template if
applicable.
- FIXME: this might be a good place to go through the [Personas and Pathways](
https://mozilla.github.io/open-leadership-training-series/articles/building-communities-of-contributors/bring-on-contributors-using-personas-and-pathways/) 
exercise, focusing on user and contributor personas, then add something to
CONTRIBUTING.md about the target user of the package.

6. Create a NEWS file to your repository about changes you make to it:
    - Create a file called `NEWS.md` in the top level of your project by running
    `usethis::use_news_md()`.
    - Review the [Tidyverse NEWS style guide](https://style.tidyverse.org/news.html)
    on best practices for writing in your NEWS file.

7. Create a reprex to practice creating Reproducible Examples for debugging and
getting help:
    - Use the `reprex` package to create a minimal reproducible example for one
    of the functions in your package.
    - Add your process for using the reprex, and what you've learned from it,
    into the `CONTRIBUTING.md` file, as an example of how to contribute
    and use issues.

### Exercises for second project assignment
