# Workflows when using GitHub {#github-workflows}

<!-- Brief summary of list of actions/code/concepts done -->

- Exercises: Conceptual, how can people effectively collaborate on a project?
    - Maybe get them to create a vignette article with their answers/thoughts? 
    Articles that are ignored from package build. (we discussed this a bit
    already, needs more discussion)
- Branching/forks
    - Text: what is it, power of it, features.
- Pull Requests
    - Exercises: Reading, comprehension, and conceptual understanding (non-coding)
- Security:
    - Text: why its important, how to do it.
    - Code:
        - usethis::git_sitrep()
        - usethis::gh_token_help()
        - usethis::create_github_token()
    - Text: Save PAT in password manager (they should have this anyway, but this forces them)
    - Code: 
        - gitcreds::gitcreds_set(), everytime computer restarts, only saved for session
    - Exercises: Conceptual understanding of PAT, tokens, and password managers (non-coding)
    - Exercises: Setting up PAT, saving PAT, using gitcreds at start of session
- usethis pr_* helpers
    - Text: Why use these over GitHub interface
- As contributor/sole developer/as team
    - usethis::pr_init()
        - Make changes to Contributing about using pr helpers
    - usethis::pr_push()
    - usethis::pr_view()
    - usethis::pr_pause()
    - Exercise: Make an edit on the PR changed files using GitHubs edit interface.
    - Exercise: Create another PR for adding ...FIXME.
    - usethis::pr_resume()
    - usethis::pr_pull()
    - usethis::pr_merge_main()
    - usethis::pr_push()
    - PR has been merged
    - gert::git_branch_list() # Show list of created branches
    - usethis::pr_finish()
    - Exercise: Do the same process for the other PR you had created.
    - gert::git_branch_list() # Show list of created branches
    - usethis::pr_init()
    - usethis::pr_forget()
    - gert::git_branch_list() # Show list of created branches
- As maintainer, reviewer, team lead. (Hard to do exercises for this... need to
think how this part would be implemented/practiced.)
    - usethis::pr_view()
    - usethis::pr_fetch()
    - usethis::pr_push()
    - usethis::pr_merge_main()
    - usethis::pr_view()
    - Manually merge PR.
    - usethis::pr_finish()
    - Exercise: Conceptual, write out/describe how this process could improve
    collaboration and teamwork.

## Learning objectives {#11-learning-objectives}

- Learn what branching in Git is, about the concept behind it, and why it's so
powerful as a software development tool.
- Learn how to use branching in Git, as well as the related concept of forking
and Pull Requests in GitHub, by making use of usethis' Pull Request helper
functions.
- Understand how the concepts and workflows shown could be applied in team
or contributor settings and why they are powerful tools for collaboration and
teamwork.

## Collaborating with Git

### Exercise: Consider ways of collaborating and their effectiveness.

TODO: Taken from PyRSE, needs modification.

Chartreuse and Fuchsia are collaborating on a project.
Describe what is in each of the four repositories involved after each of the steps below.

1.  Chartreuse creates a repository containing a `README.md` file on GitHub
    and clones it to their desktop.
2.  Fuchsia forks that repository on GitHub
    and clones their copy to their desktop.
3.  Fuchsia adds a file `fuchsia.txt` to the `main` branch of their desktop repository
    and pushes that change to their repository on GitHub.
4.  Fuchsia creates a pull request from the `main` branch of their repository on GitHub
    to the `main` branch of Chartreuse's repository on GitHub.
5.  Chartreuse does *not* merge Fuchsia's PR.
    Instead,
    they add a file `chartreuse.txt` to the `main` branch of their desktop repository
    and push that change to their repository on GitHub.
6.  Fuchsia adds a remote to their desktop repository called `upstream`
    that points at Chartreuse's repository on GitHub
    and runs `git pull upstream main`,
    then merges any changes or conflicts.
7.  Fuchsia pushes from the `main` branch of their desktop repository
    to the `main` branch of their GitHub repository.
8.  Chartreuse merges Fuchsia's pull request.
9.  Chartreuse runs `git pull origin main` on the desktop.

### Branches and forks

- when to use either

- How to use them effectively
    - How branches can help both solo and as a team
    - When forks are good vs being contributor on a repo (write access)
    
## Collaborating with GitHub

- What are PRs, how to use them (either non-fork or fork)
    - Keeping things targeted (e.g. a PR should conceptually
    fix or add similarly related things)
    
### Exercise: Reading and comprehension task?

### Exercise: Do these situations warrant a branch or fork?

Concept and thought, not code based.

1. 
2. 
3. 

### Exercise: Consider some situations that would benefit from using PRs

Concept and thought, not code based.

## Securing your GitHub

Why this is important, etc.

- usethis::git_sitrep()

- usethis::gh_token_help()

- usethis::create_github_token()

- Save PAT in password manager (they should have this anyway, but this forces them)
- gitcreds::gitcreds_set(), everytime computer restarts, only saved for session
    - Exercises: Conceptual understanding of PAT, tokens, and password managers (non-coding)
    - Exercises: Setting up PAT, saving PAT, using gitcreds at start of session

## Simplified collaboration with usethis

### Workflows for contributors on solo or team projects

- usethis pr_* helpers
- As contributor/sole developer/as team
    - usethis::pr_init()
        - Make changes to Contributing about using pr helpers
    - usethis::pr_push()
    - usethis::pr_view()
    - usethis::pr_pause()
    - Exercise: Make an edit on the PR changed files using GitHubs edit interface.
    - Exercise: Create another PR for adding ...FIXME.
    - usethis::pr_resume()
    - usethis::pr_pull()
    - usethis::pr_merge_main()
    - usethis::pr_push()
    - PR has been merged
    - gert::git_branch_list() # Show list of created branches
    - usethis::pr_finish()
    - Exercise: Do the same process for the other PR you had created.

## Removing a branch

- gert::git_branch_list() # Show list of created branches
- usethis::pr_init()
- usethis::pr_forget()
- gert::git_branch_list() # Show list of created branches

### Exercise: Deleting a branch without merging

TODO: Taken from PyRSE, needs more modifying.

Create a branch called `experiment` with pr_init(). Create a new vignette
called `experiment.Rmd` with use_article(). Add your name to the YAML header.
Commit the changes and move back to main branch with pr_pause().

1.  Run pr_forget()...

### Workflows for project maintainers or team leads

- As maintainer, reviewer, team lead. (Hard to do exercises for this... need to
think how this part would be implemented/practiced.)
    - usethis::pr_view()
    - usethis::pr_fetch()
    - usethis::pr_push()
    - usethis::pr_merge_main()
    - usethis::pr_view()
    - Manually merge PR.
    - usethis::pr_finish()
    - Exercise: Conceptual, write out/describe how this process could improve
    collaboration and teamwork.
    
```{r, eval=FALSE}
# NOTE: This is some of the code used throughout the chapter, put in a format
# that any changes to the project can be automatically added at final book build.
library(usethis)

# create_github_token()
# gitcreds::
# TODO: Talk about security?
# gh_token_help()

# TODO: Add to CONTRIBUTING guidelines about using usethis pr
pr_init("contributing-update")

# Not added to code, but added to text. Using this so package gets built exactly
# as stated in content.
# contributing_text <- readLines(here::here("CONTRIBUTING.md")) # Yea?
contribute_pr_helpers_text <- "
TODO: Add text about using usethis
"
# writeLines(contribute_pr_helpers_text, here::here("CONTRIBUTING.md"))

# gert::git_add("CONTRIBUTING.md")
# gert::git_commit("add to use usethis pr_ helpers in contributing")

# pr_push()
# Go into GitHub and merge PR (R function to do that? pr_fetch? git_merge then
# git_push?)

# pr_init("another-branch") # TODO: other changes?
# pr_pause()
# Someone reviews the PR and accepts it
# pr_resume("another-branch")
# 
# pr_finish()
```

## Final exercise {#11-final-exercise}

Complete these tasks as part of the package for your first project assignment.
Refer to the project assignment TODO: Add ref to assignment here?
for more details. As with the other final exercises,
we will use the `kenyaweather` package as an example and reference.
Don't forget to add and commit any changes you've made into the Git history
after each task, and pushing to GitHub.

TODO: Add guidelines about using usethis::pr_ helpers? or at least link to it?

1. Create a new Issue on your package's GitHub repository about an idea for a
function or to fix an existing function (you don't have to actually fix it).
2. Using the `pr_*` helpers, create a new branch called `fixing-function` using
`pr_init("fixing-function")`.
    - Make the following changes to the FIXME function:
        - FIXME: Not really sure what to say here. Could indicate that instructor has more detailed instructions. OR maybe not. Let's come back to this.
    - Add and commit these changes.
5. Push the changes up to GitHub by using `pr_push()`.
    - Create a new Pull Request based on the push in the newly opened up GitHub
    page (this should open automatically with `pr_push()`).
    - Merge the newly created Pull Request.
6. In your local Git repository, run `pr_finish()` to finish up.

Repeat this process one more time by making changes to:

- FIXME: Adding to README maybe or one of the vignettes?

```{r, eval=FALSE}
# This is the actions done to the project package
# 1. new issue
# githubr::gh_new_issue ? or gh package?
# new_number <- githubr::gh_issue_list ?

# 2. 
# pr_init("fixing-function")

# 3. 
# TODO: What to change?

# 4.
# git_add("R/TODO")
# git_commit(str_c("update function to have TODO. Closes ", new_number))

# 5.
# pr_push()
# pr_pause()
# git_merge("fixing-function")
# git_push() ?
# pr_resume("fixing-function")
# pr_finish()
```

### Exercises for second project assignment?
